---
title: "D34.Pacemaker.Cells"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(RColorBrewer)
library(scDblFinder)
library(SummarizedExperiment)
library(stringr)
library(tidyr)
library(ggforce)
library(cluster)
library(rBCS)
library(reshape2)
library(ggpubr)
library(scCustomize)
library(ComplexHeatmap)
library(circlize)
library(viridis)
library(paletteer)
```


```{r}
load("/Data/iPSC_pacemaker/Seurat/D34_First/D34.filtered.samples.RData")
pacemarker.cells <- subset(filtered.samples, subset = orig.ident == "D34_r0_P" | orig.ident == "D34_r1_P" | orig.ident == "D34_r2_P")
pacemarker.cells
#An object of class Seurat 
#27519 features across 20712 samples within 1 assay 
#Active assay: RNA (27519 features, 0 variable features)
```

```{r}
# Normalize the counts
seurat_phase <- NormalizeData(pacemarker.cells)
```

Evaluating effects of cell cycle.
```{r}
# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

# Score cells for cell cycle
seurat_phase <- CellCycleScoring(seurat_phase, 
                                 g2m.features = g2m.genes, 
                                 s.features = s.genes)

## Error
#Warning: The following features are not present in the object: MLF1IP, not searching for symbol synonyms
#Warning: The following features are not present in the object: FAM64A, HN1, not searching for symbol synonyms

# Manually check the ENSG ID for these three genes and found that:
## MLF1IP is CENPU in our reference (ENSG00000151725)
## FAM64A is PIMREG in our reference (ENSG00000129195)
## HN1 is JPT1 in our reference (ENSG00000189159)

intersect(c("FAM64A", "HN1", "MLF1IP"), s.genes)
#[1] "MLF1IP"
intersect(c("FAM64A", "HN1", "MLF1IP"), g2m.genes)
#[1] "FAM64A" "HN1"

# Manually change the names of these cell cycle genes.
s.genes[s.genes=="MLF1IP"] = "CENPU"
g2m.genes[g2m.genes=="FAM64A"] = "PIMREG"
g2m.genes[g2m.genes=="HN1"] = "JPT1"

# Redo score cells for cell cycle
seurat_phase <- CellCycleScoring(seurat_phase, 
                                 g2m.features = g2m.genes, 
                                 s.features = s.genes)
# Calculate the percentage of cells in different phases.
phase_count <- table(seurat_phase@meta.data$orig.ident, seurat_phase@meta.data$Phase)
phase_per <- prop.table(table(seurat_phase@meta.data$orig.ident, seurat_phase@meta.data$Phase), 1)*100
write.table(phase_count, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/CellCyclePhaseCount.txt", quote = F, sep = "\t")
write.table(phase_per, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/CellCyclePhasePercentage.txt", quote = F, sep = "\t")

# Visualize the distribution of cell cycle markers across
RidgePlot(seurat_phase, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), group.by = "Phase", ncol = 2) & scale_fill_manual(values = custom_colors$discrete)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/Cell.cycle.markers.by.phase.pdf")

RidgePlot(seurat_phase, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), group.by = "orig.ident", ncol = 2) & scale_y_discrete(limits = rev(c("D34_r0_P", "D34_r1_P", "D34_r2_P"))) & scale_fill_manual(values = custom_colors$discrete)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/Cell.cycle.markers.by.sample.pdf")
```

```{r}
# Identify the most variable genes
seurat_phase <- FindVariableFeatures(seurat_phase, 
                     selection.method = "vst",
                     nfeatures = 4000, 
                     verbose = FALSE)
		     
# Scale the counts
seurat_phase <- ScaleData(seurat_phase)

# Perform PCA
seurat_phase <- RunPCA(seurat_phase, ndims.print = 1:40, nfeatures.print = 10)

# Run UMAP
seurat_phase <- RunUMAP(seurat_phase, dims = 1:40)

# Plot the UMAP colored by cell cycle phase
DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "Phase",
        split.by = "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/SplitCellCyclePlot.pdf")

DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/MergedCellCyclePlot.pdf")

DimPlot(seurat_phase,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        group.by= "orig.ident") + scale_color_discrete(limits = c("D34_r0_P", "D34_r1_P", "D34_r2_P"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/MergedCellCyclePlot_Samples.pdf")

#DimHeatmap(seurat_phase, dims = c(8, 10))
```
We do see large differences due to cell cycle. Based on this plot, we would regress out the variation due to cell cycle.

Evaluating effects of mitochodrial expression.
```{r}
# Check quartile values
summary(seurat_phase@meta.data$percent.mt)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#   0.00   11.19   15.59   16.49   20.65   49.96
# Turn mitoRatio into categorical factor vector based on quartile values
seurat_phase@meta.data$mitoFr <- cut(seurat_phase@meta.data$percent.mt, 
                   breaks=c(-Inf, 11.19, 15.59, 20.65, Inf), 
                   labels=c("Low","Medium","Medium high", "High"))

# Plot the PCA colored by mitoFr
DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "mitoFr",
        split.by = "mitoFr")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/SplitMitoFrPlot.pdf")

DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "mitoFr")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/MergedMitoFrPlot.pdf")
```
We do not see large differences due to mitochondrial percentage for the same Tissue. Based on this plot, we would not regress out the variation due to mitochondrial percentage.

Add cell cycle score to pacemarker.cells for SCTransform.
```{r}
table(rownames(seurat_phase@meta.data)==rownames(pacemarker.cells@meta.data))
pacemarker.cells@meta.data <- mutate(pacemarker.cells@meta.data, S.Score = seurat_phase@meta.data$S.Score, G2M.Score = seurat_phase@meta.data$G2M.Score, Phase = seurat_phase@meta.data$Phase)
```

Apply sctransform normalization while regress out cell cycle scoring.
```{r}
# SCTranform
## Adjust the limit for allowable object sizes within R (Default is 500 * 1024 ^ 2 = 500 Mb) using the following code:
#options(future.globals.maxSize = 10000 * 1024^2)
pacemarker.cells <- SCTransform(pacemarker.cells, method = "glmGamPoi", vars.to.regress = c("G2M.Score", "S.Score"), variable.features.n = 5000)

# Save the seurat object
saveRDS(pacemarker.cells, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/SCTransform.pacemarker.cells.rds")

#filtered.samples <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/SCTransform.filtered.samples.rds")
```


```{r}
pacemarker.cells <- RunPCA(pacemarker.cells)
DimPlot(pacemarker.cells, reduction = "pca", group.by = "orig.ident", cols = custom_colors$discrete) + scale_color_discrete(limits = c("D34_r0_P", "D34_r1_P", "D34_r2_P"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/PCs/MergedPCAPlot.pdf")

pacemarker.cells$orig.ident <- factor(pacemarker.cells$orig.ident, levels = unique(pacemarker.cells$orig.ident))

DimPlot(pacemarker.cells, reduction = "pca", split.by = "orig.ident", group.by = "orig.ident", cols = custom_colors$discrete) + scale_color_discrete(limits = c("D34_r0_P", "D34_r1_P", "D34_r2_P"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/PCs/SplitPCAPlot.pdf", width = 30, height = 5)
```

Determining how many PCs to include in the clustering step to ensure that we are capturing the majority of the variation, or cell types, present in our dataset.
```{r}
# Explore heatmap of PCs
pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/PCs/HeatmapOfPCs_1_12.pdf", width = 25, height = 15)
#DimHeatmap(pacemarker.cells, dims = 1:25, cells = 500, balanced = TRUE, fast = FALSE)
DimHeatmap(pacemarker.cells, dims = 1:12, cells = 500, balanced = TRUE)
dev.off()

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/PCs/HeatmapOfPCs_13_24.pdf", width = 25, height = 15)
DimHeatmap(pacemarker.cells, dims = 13:24, cells = 500, balanced = TRUE)
dev.off()

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/PCs/HeatmapOfPCs_25_36.pdf", width = 25, height = 15)
DimHeatmap(pacemarker.cells, dims = 25:36, cells = 500, balanced = TRUE)
dev.off()

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/PCs/HeatmapOfPCs_37_48.pdf", width = 25, height = 15)
DimHeatmap(pacemarker.cells, dims = 37:48, cells = 500, balanced = TRUE)
dev.off()

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/PCs/HeatmapOfPCs_49_50.pdf", width = 25, height = 15)
DimHeatmap(pacemarker.cells, dims = 49:50, cells = 500, balanced = TRUE)
dev.off()

# Plot the elbow plot
ElbowPlot(object = pacemarker.cells, ndims = 50)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/PCs/ElbowPlot.pdf")
```
Decision: Use the first 40 PCs to generate the clusters.

Cluster the cells
```{r}
# Determine the K-nearest neighbor graph
pacemarker.cells <- FindNeighbors(object = pacemarker.cells, 
                                dims = 1:40)
                                
# Determine the clusters for various resolutions                                
pacemarker.cells <- FindClusters(object = pacemarker.cells,
                               resolution = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

# Look at cluster IDs of the first 5 cells
head(Idents(pacemarker.cells), 5)

#pacemarker.cells@meta.data <- select(pacemarker.cells@meta.data, c(1,25,2:10,26:34,11:24))


# Count the number of clusters at different resolutions.
Ncluster <- c()
for(res in c(12:33)){
  Ncluster <- c(Ncluster, length(unique(as.vector(pacemarker.cells@meta.data[,res]))))
#unique(pacemarker.cells@meta.data$SCT_snn_res.2)
}
res_ncluster <- data.frame(Res = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0), Ncluster = Ncluster)
write.table(res_ncluster,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/ResolutionVsNclusters.txt", quote = F, sep = "\t", row.names = F)
```

Run non-linear dimensional reduction (UMAP/tSNE)
```{r}
pacemarker.cells <- RunUMAP(pacemarker.cells, dims = 1:40)

# Plot the UMAP
# Assign identity of clusters
Idents(object = pacemarker.cells) <- "SCT_snn_res.0.6"
DimPlot(pacemarker.cells,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/UMAP.res.0.6.pdf", width = 10, height = 8)
```
Decision: Use resolution 0.6.

Segregation of clusters by sample
```{r}
# Extract identity and sample information from seurat object to determine the number of cells per cluster per sample
n_cells <- FetchData(pacemarker.cells, 
                     vars = c("ident", "orig.ident")) %>%
        dplyr::count(ident, orig.ident) %>%
        tidyr::spread(ident, n)

write.table(n_cells,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/N_Cells_For_Each_Clusters.txt", quote = F, sep = "\t", row.names = F)

# UMAP of cells in each cluster by sample
DimPlot(pacemarker.cells, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/SplitSamples.UMAP.res.0.6.pdf", width = 15, height = 5)

DimPlot(pacemarker.cells, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/MergedSamples.UMAP.res.0.6.pdf", width = 10, height = 8)

DimPlot(pacemarker.cells, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.pdf", width = 10, height = 8)

pacemarker.cells$Tissue <- factor(pacemarker.cells$Tissue, levels = unique(pacemarker.cells$Tissue))
DimPlot(pacemarker.cells, 
        label = TRUE, 
        split.by = "Tissue",
        group.by = "Tissue")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/Split.Tissues.UMAP.res.0.6.pdf", width = 10, height = 5)

DimPlot(pacemarker.cells, 
        label = TRUE, 
        #cols = brewer.pal(6,"Dark2"),
        group.by = "Tissue")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/Merged.Tissues.UMAP.res.0.6.pdf", width = 10, height = 8)

```
Using the top 5000 variable genes among D34 pacemaker cells shows batch effect. This is inconsistent with the analysis using all the D34 cells including pacemaker, atrial and ventricular cells, which also used top 5000 variable genes. However, these two set of 5000 variable genes are different. The reason is that the genes leading to batch effect in D34 pacemaker cell clustering is not included in the D34 all cell type clustering as difference among cell types dominate the difference between batches in D34 pacemaker cells.


Test to use top 3000 genes
```{r}
pacemarker.cells.3000 <- SCTransform(pacemarker.cells, method = "glmGamPoi", vars.to.regress = c("G2M.Score", "S.Score"), variable.features.n = 3000)

pacemarker.cells.3000 <- RunPCA(pacemarker.cells.3000)

# Determine the K-nearest neighbor graph
pacemarker.cells.3000 <- FindNeighbors(object = pacemarker.cells.3000, 
                                dims = 1:40)
                                
# Determine the clusters for various resolutions                                
pacemarker.cells.3000 <- FindClusters(object = pacemarker.cells.3000,
                               resolution = c(0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

pacemarker.cells.3000 <- RunUMAP(pacemarker.cells.3000, dims = 1:40)

# Plot the UMAP
# Assign identity of clusters
Idents(object = pacemarker.cells.3000) <- "SCT_snn_res.0.6"

# UMAP of cells in each cluster by sample
DimPlot(pacemarker.cells.3000, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/SplitSamples.UMAP.res.0.6.3000.pdf", width = 15, height = 5)

DimPlot(pacemarker.cells.3000, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/MergedSamples.UMAP.res.0.6.3000.pdf", width = 10, height = 8)

DimPlot(pacemarker.cells.3000, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.3000.pdf", width = 10, height = 8)

length(intersect(rownames(filtered.D34.samples[["SCT"]]@scale.data), rownames(pacemarker.cells.3000[["SCT"]]@scale.data)))
#[1] 2762
```
Less batch effect was observed using top 3000 genes.


Test to see whether using the same genes in the D34 all cell type clustering generate patterns without batch effect (3643 overlapping genes between the two top 5000 variable genes as shown in below analysis)
```{r}
filtered.D34.samples <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/Clustered.filtered.samples.rds")

dim(filtered.D34.samples[["SCT"]]@scale.data)
#[1]  5000 64322
dim(filtered.D34.samples[["SCT"]]@data)
#[1] 27519 64322
dim(filtered.D34.samples[["SCT"]]@counts)
#[1] 27519 64322
write.table(rownames(filtered.D34.samples[["SCT"]]@scale.data),"/Data/iPSC_pacemaker/Seurat/D34_First/SCT_Top5000genes.txt", sep = "\t", quote = F, row.names = F, col.names = F)

length(intersect(rownames(filtered.D34.samples[["SCT"]]@scale.data), rownames(pacemarker.cells[["SCT"]]@scale.data)))
#[1] 3643

pacemarker.cells.all.D34.cells.top.5000 <- subset(filtered.samples, features = rownames(filtered.D34.samples[["SCT"]]@scale.data))
length(intersect(rownames(filtered.D34.samples[["SCT"]]@scale.data), rownames(pacemarker.cells.all.D34.cells.top.5000[["RNA"]]@counts)))
#[1] 5000

pacemarker.cells.all.D34.cells.top.5000 <- subset(pacemarker.cells.all.D34.cells.top.5000, subset = orig.ident == "D34_r0_P" | orig.ident == "D34_r1_P" | orig.ident == "D34_r2_P")
```


```{r}
# Normalize the counts
seurat_phase <- NormalizeData(pacemarker.cells.all.D34.cells.top.5000)
```

Evaluating effects of cell cycle.
```{r}
# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

# Score cells for cell cycle
seurat_phase <- CellCycleScoring(seurat_phase, 
                                 g2m.features = g2m.genes, 
                                 s.features = s.genes)

## Error
#Warning: The following features are not present in the object: MLF1IP, not searching for symbol synonyms
#Warning: The following features are not present in the object: FAM64A, HN1, not searching for symbol synonyms

# Manually check the ENSG ID for these three genes and found that:
## MLF1IP is CENPU in our reference (ENSG00000151725)
## FAM64A is PIMREG in our reference (ENSG00000129195)
## HN1 is JPT1 in our reference (ENSG00000189159)

intersect(c("FAM64A", "HN1", "MLF1IP"), s.genes)
#[1] "MLF1IP"
intersect(c("FAM64A", "HN1", "MLF1IP"), g2m.genes)
#[1] "FAM64A" "HN1"

# Manually change the names of these cell cycle genes.
s.genes[s.genes=="MLF1IP"] = "CENPU"
g2m.genes[g2m.genes=="FAM64A"] = "PIMREG"
g2m.genes[g2m.genes=="HN1"] = "JPT1"

# Redo score cells for cell cycle
seurat_phase <- CellCycleScoring(seurat_phase, 
                                 g2m.features = g2m.genes, 
                                 s.features = s.genes)
# Calculate the percentage of cells in different phases.
phase_count <- table(seurat_phase@meta.data$orig.ident, seurat_phase@meta.data$Phase)
phase_per <- prop.table(table(seurat_phase@meta.data$orig.ident, seurat_phase@meta.data$Phase), 1)*100
#write.table(phase_count, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/CellCyclePhaseCount.txt", quote = F, sep = "\t")
#write.table(phase_per, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/CellCyclePhasePercentage.txt", quote = F, sep = "\t")

# Visualize the distribution of cell cycle markers across
#RidgePlot(seurat_phase, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), group.by = "Phase", ncol = 2) & scale_fill_manual(values = custom_colors$discrete)
#ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/Cell.cycle.markers.by.phase.pdf")

#RidgePlot(seurat_phase, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), group.by = "orig.ident", ncol = 2) & scale_y_discrete(limits = rev(c("D34_r0_P", "D34_r1_P", "D34_r2_P"))) & scale_fill_manual(values = custom_colors$discrete)
#ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/Cell.cycle.markers.by.sample.pdf")
```

```{r}
# Identify the most variable genes
seurat_phase <- FindVariableFeatures(seurat_phase, 
                     selection.method = "vst",
                     nfeatures = 4000, 
                     verbose = FALSE)
		     
# Scale the counts
seurat_phase <- ScaleData(seurat_phase)

# Perform PCA
seurat_phase <- RunPCA(seurat_phase, ndims.print = 1:40, nfeatures.print = 10)

# Run UMAP
seurat_phase <- RunUMAP(seurat_phase, dims = 1:40)

# Plot the UMAP colored by cell cycle phase
DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "Phase",
        split.by = "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/SplitCellCyclePlot.D34.cells.top.5000.pdf")

DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/MergedCellCyclePlot.D34.cells.top.5000.pdf")

DimPlot(seurat_phase,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        group.by= "orig.ident") + scale_color_discrete(limits = c("D34_r0_P", "D34_r1_P", "D34_r2_P"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Covariates/MergedCellCyclePlot_Samples.D34.cells.top.5000.pdf")

#DimHeatmap(seurat_phase, dims = c(8, 10))
```
We do see large differences due to cell cycle. Based on this plot, we would regress out the variation due to cell cycle.


Add cell cycle score to pacemarker.cells.all.D34.cells.top.5000 for SCTransform.
```{r}
table(rownames(seurat_phase@meta.data)==rownames(pacemarker.cells.all.D34.cells.top.5000@meta.data))
pacemarker.cells.all.D34.cells.top.5000@meta.data <- mutate(pacemarker.cells.all.D34.cells.top.5000@meta.data, S.Score = seurat_phase@meta.data$S.Score, G2M.Score = seurat_phase@meta.data$G2M.Score, Phase = seurat_phase@meta.data$Phase)
```

Apply sctransform normalization while regress out cell cycle scoring.
```{r}
# SCTranform
## Adjust the limit for allowable object sizes within R (Default is 500 * 1024 ^ 2 = 500 Mb) using the following code:
#options(future.globals.maxSize = 10000 * 1024^2)
pacemarker.cells.all.D34.cells.top.5000 <- SCTransform(pacemarker.cells.all.D34.cells.top.5000, method = "glmGamPoi", vars.to.regress = c("G2M.Score", "S.Score"), variable.features.n = 5000)

# Save the seurat object
#saveRDS(pacemarker.cells.all.D34.cells.top.5000, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/SCTransform.pacemarker.cells.all.D34.cells.top.5000.rds")

#filtered.samples <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/SCTransform.filtered.samples.rds")
```


```{r}
pacemarker.cells.all.D34.cells.top.5000 <- RunPCA(pacemarker.cells.all.D34.cells.top.5000)
```

Cluster the cells
```{r}
# Determine the K-nearest neighbor graph
pacemarker.cells.all.D34.cells.top.5000 <- FindNeighbors(object = pacemarker.cells.all.D34.cells.top.5000, 
                                dims = 1:40)
                                
# Determine the clusters for various resolutions                                
pacemarker.cells.all.D34.cells.top.5000 <- FindClusters(object = pacemarker.cells.all.D34.cells.top.5000,
                               resolution = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))
```

Run non-linear dimensional reduction (UMAP/tSNE)
```{r}
pacemarker.cells.all.D34.cells.top.5000 <- RunUMAP(pacemarker.cells.all.D34.cells.top.5000, dims = 1:40)

# Plot the UMAP
# Assign identity of clusters
Idents(object = pacemarker.cells.all.D34.cells.top.5000) <- "SCT_snn_res.0.6"
DimPlot(pacemarker.cells.all.D34.cells.top.5000,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/UMAP.res.0.6.D34.cells.top.5000.pdf", width = 10, height = 8)
```
Decision: Use resolution 0.6.

Segregation of clusters by sample
```{r}
# UMAP of cells in each cluster by sample
DimPlot(pacemarker.cells.all.D34.cells.top.5000, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/SplitSamples.UMAP.res.0.6.D34.cells.top.5000.pdf", width = 15, height = 5)

DimPlot(pacemarker.cells.all.D34.cells.top.5000, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/MergedSamples.UMAP.res.0.6.D34.cells.top.5000.pdf", width = 10, height = 8)

DimPlot(pacemarker.cells.all.D34.cells.top.5000, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.D34.cells.top.5000.pdf", width = 10, height = 8)
```
Much less clear batch effect is seen with the top 5000 variable genes in all D34 cell types.




#####################################################################################################################
#####################################################################################################################
#####################################################################################################################
Decide to do integration for the 3 replicates to see if batch effect is still observed using top 3000 variable genes.
```{r}
# split the dataset into a list of three seurat objects (each for one sample)
pacemaker.list <- SplitObject(pacemarker.cells, split.by = "orig.ident")

# normalize and identify variable features for each dataset independently
pacemaker.list <- lapply(X = pacemaker.list, FUN = function(x) {
    x <- SCTransform(x, method = "glmGamPoi", vars.to.regress = c("G2M.Score", "S.Score"), variable.features.n = 5000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = pacemaker.list, nfeatures = 3000)
pacemaker.list <- PrepSCTIntegration(object.list = pacemaker.list, anchor.features = features)

# Perform integration
pacemaker.anchors <- FindIntegrationAnchors(object.list = pacemaker.list, normalization.method = "SCT", anchor.features = features)
pacemaker.combined.sct <- IntegrateData(anchorset = pacemaker.anchors, normalization.method = "SCT")

pacemaker.combined.sct <- RunPCA(pacemaker.combined.sct, verbose = FALSE)
pacemaker.combined.sct <- RunUMAP(pacemaker.combined.sct, reduction = "pca", dims = 1:40)

pacemaker.combined.sct <- FindNeighbors(pacemaker.combined.sct, reduction = "pca", dims = 1:40)
pacemaker.combined.sct <- FindClusters(pacemaker.combined.sct, resolution = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

Idents(object = pacemaker.combined.sct) <- "integrated_snn_res.0.6"
DimPlot(pacemaker.combined.sct,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/UMAP.res.0.6.D34.cells.3000.features.pdf", width = 10, height = 8)

#Segregation of clusters by sample
# UMAP of cells in each cluster by sample
DimPlot(pacemaker.combined.sct, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/SplitSamples.UMAP.res.0.6.D34.cells.top.3000.features.pdf", width = 15, height = 5)

DimPlot(pacemaker.combined.sct, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/MergedSamples.UMAP.res.0.6.D34.cells.top.3000.features.pdf", width = 10, height = 8)

DimPlot(pacemaker.combined.sct, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.D34.cells.top.3000.features.pdf", width = 10, height = 8)
```

Try using 5000 genes.
```{r}
features <- SelectIntegrationFeatures(object.list = pacemaker.list, nfeatures = 5000)
pacemaker.list <- PrepSCTIntegration(object.list = pacemaker.list, anchor.features = features)

# Perform integration
pacemaker.anchors <- FindIntegrationAnchors(object.list = pacemaker.list, normalization.method = "SCT", anchor.features = features)
pacemaker.combined.sct <- IntegrateData(anchorset = pacemaker.anchors, normalization.method = "SCT")

pacemaker.combined.sct <- RunPCA(pacemaker.combined.sct, verbose = FALSE)
pacemaker.combined.sct <- RunUMAP(pacemaker.combined.sct, reduction = "pca", dims = 1:40)

pacemaker.combined.sct <- FindNeighbors(pacemaker.combined.sct, reduction = "pca", dims = 1:40)
pacemaker.combined.sct <- FindClusters(pacemaker.combined.sct, resolution = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

Idents(object = pacemaker.combined.sct) <- "integrated_snn_res.0.6"
DimPlot(pacemaker.combined.sct,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/UMAP.res.0.6.D34.cells.5000.features.pdf", width = 10, height = 8)

#Segregation of clusters by sample
# UMAP of cells in each cluster by sample
DimPlot(pacemaker.combined.sct, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/SplitSamples.UMAP.res.0.6.D34.cells.top.5000.features.pdf", width = 15, height = 5)

DimPlot(pacemaker.combined.sct, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/MergedSamples.UMAP.res.0.6.D34.cells.top.5000.features.pdf", width = 10, height = 8)

DimPlot(pacemaker.combined.sct, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.D34.cells.top.5000.features.pdf", width = 10, height = 8)

DimPlot(pacemaker.combined.sct, 
        label = FALSE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.D34.cells.top.5000.features.no.legend.pdf", width = 10, height = 8)
```
The pattern stayed similar. Decided to use 5000 genes for downstream.


```{r}
# Count the number of clusters at different resolutions.
Ncluster <- c()
for(res in c(11:33)){
  Ncluster <- c(Ncluster, length(unique(as.vector(pacemaker.combined.sct@meta.data[,res]))))
#unique(pacemarker.cells@meta.data$SCT_snn_res.2)
}
res_ncluster <- data.frame(Res = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0), Ncluster = Ncluster)
write.table(res_ncluster,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/ResolutionVsNclusters.txt", quote = F, sep = "\t", row.names = F)
```


Segregation of clusters by cell cycle phase
```{r}
# Explore whether clusters segregate by cell cycle phase
DimPlot(pacemaker.combined.sct,
        group.by = "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Merged.CellCycle.UMAP.res.0.6.pdf", width = 10, height = 8)
DimPlot(pacemaker.combined.sct,
        group.by = "Phase",
        split.by = "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Split.CellCycle.UMAP.res.0.6.pdf", width = 25, height = 10)
```


Segregation of clusters by various sources of uninteresting variation
```{r}
# Determine metrics to plot present in seurat_integrated@meta.data
metrics <-  c("nCount_RNA", "nFeature_RNA", "S.Score", "G2M.Score", "percent.mt")

FeaturePlot(pacemaker.combined.sct, 
            reduction = "umap", 
            features = metrics,
            pt.size = 0.4, 
            order = TRUE,
            #min.cutoff = 'q10',
            label = TRUE)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Metrics.UMAP.res.0.6.pdf", width = 10, height = 10)
```


```{r}
# Save the object at this point so that it can easily be loaded back in without having to rerun the computationally intensive steps performed above, or easily shared with collaborators.
saveRDS(pacemaker.combined.sct, file = "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Clustered.pacemaker.combined.sct.rds")

#pacemaker.combined.sct <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Clustered.pacemaker.combined.sct.rds")
```

Exploring known cell type markers
```{r}
# Pacemaker cell genes
FeaturePlot(pacemaker.combined.sct, 
            reduction = "umap", 
            features = c("TBX18", "SHOX2", "ISL1", "TBX3"), 
            order = FALSE,
            #min.cutoff = 'q10', 
            label = TRUE)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/PacemakerCell_Genes.UMAP.res.0.6.pdf", width = 12, height = 10)

# Violin plot
VlnPlot(pacemaker.combined.sct, c("TBX18", "SHOX2", "ISL1", "TBX3"), group.by = "orig.ident", pt.size = 0, cols = custom_colors$discrete, ncol = 4)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/PacemakerCell_Genes.Violin.UMAP.res.0.6.pdf", width = 16, height = 5)
```

```{r}
DotPlot(pacemaker.combined.sct, features = rev(c("NKX2-5", "TNNT2", "MYH6", "MYH7","TBX18", "SHOX2", "ISL1", "TBX3", "HCN4", "KCNJ3", "GJD3", "MYL2", "HEY2", "IRX4", "MYL7", "NPPA", "NR2F2", "MSX2", "TBX2")), cols= "Spectral", group.by = "orig.ident") + RotatedAxis() + coord_flip()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/James_Markers_DotPlot.UMAP.res.0.6.pdf", width = 8, height = 6)
```

```{r}
# Single cell heatmap of feature expression
DoHeatmap(pacemaker.combined.sct, features = c("NKX2-5", "TNNT2", "MYH6", "MYH7","TBX18", "SHOX2", "ISL1", "TBX3", "HCN4", "KCNJ3", "GJD3", "MYL2", "HEY2", "IRX4", "MYL7", "NPPA", "NR2F2", "MSX2", "TBX2"), size = 3, group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/James_Markers_Heatmap.UMAP.res.0.6.pdf")
```

Composition of samples and clusters
```{r}
table_samples_by_clusters <- pacemaker.combined.sct@meta.data %>%
  group_by(orig.ident, integrated_snn_res.0.6) %>%
  summarize(count = n()) %>%
  spread(integrated_snn_res.0.6, count, fill = 0) %>%
  ungroup() %>%
  mutate(total_cell_count = rowSums(.[c(2:ncol(.))])) %>%
  dplyr::select(c('orig.ident', 'total_cell_count', everything())) %>%
  arrange(factor(orig.ident, levels = levels(pacemaker.combined.sct@meta.data$orig.ident)))
write.table(as.data.frame(table_samples_by_clusters), "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/table_samples_by_clusters.txt", quote = F, sep = "\t")

table_clusters_by_samples <- pacemaker.combined.sct@meta.data %>%
  dplyr::rename('cluster' = 'integrated_snn_res.0.6') %>%
  group_by(cluster, orig.ident) %>%
  summarize(count = n()) %>%
  spread(orig.ident, count, fill = 0) %>%
  ungroup() %>%
  mutate(total_cell_count = rowSums(.[c(2:ncol(.))])) %>%
  select(c('cluster', 'total_cell_count', everything())) %>%
  arrange(factor(cluster, levels = levels(pacemaker.combined.sct@meta.data$integrated_snn_res.0.6)))
write.table(as.data.frame(table_clusters_by_samples), "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/table_clusters_by_samples.txt", quote = F, sep = "\t")

# Plot number of cells.
temp_labels <- pacemaker.combined.sct@meta.data %>%
  group_by(orig.ident) %>%
  tally()

p1 <- table_samples_by_clusters %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'orig.ident') %>%
  mutate(sample = factor(orig.ident, levels = unique(pacemaker.combined.sct@meta.data$orig.ident))) %>%
  ggplot(aes(sample, value)) +
  geom_bar(aes(fill = variable), position = 'stack', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = orig.ident, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 3, angle = 0
  ) +
  scale_fill_manual(name = 'Cluster', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Number of cells', labels = scales::comma, expand = c(0.01, 0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'left',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1, color = "black"),
    axis.text.y = element_text(color = "black"),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 0, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/table_samples_by_clusters_cell_number.pdf", p1, width = 8, height = 5)

temp_labels <- pacemaker.combined.sct@meta.data %>%
  group_by(integrated_snn_res.0.6) %>%
  tally() %>%
  dplyr::rename('cluster' = integrated_snn_res.0.6)

p2 <- table_clusters_by_samples %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'cluster') %>%
  mutate(cluster = factor(cluster, levels = sort(unique(pacemaker.combined.sct@meta.data$integrated_snn_res.0.6)))) %>%
  ggplot(aes(cluster, value)) +
  geom_bar(aes(fill = variable), position = 'stack', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = cluster, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Sample', values = custom_colors$discrete) +
  scale_y_continuous(labels = scales::comma, expand = c(0.01, 0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'right',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 10, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/table_clusters_by_samples_cell_number.pdf", p2, width = 10, height = 5)

# Plot percent of cells.
temp_labels <- pacemaker.combined.sct@meta.data %>%
  group_by(orig.ident) %>%
  tally()

p1 <- table_samples_by_clusters %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'orig.ident') %>%
  mutate(sample = factor(orig.ident, levels = unique(pacemaker.combined.sct@meta.data$orig.ident))) %>%
  ggplot(aes(sample, value)) +
  geom_bar(aes(fill = variable), position = 'fill', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = orig.ident, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Cluster', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Percentage [%]', labels = scales::percent_format(), expand = c(0.01,0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'left',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 0, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/table_samples_by_clusters_cell_percent.pdf", p1, width = 8, height = 5)

temp_labels <- pacemaker.combined.sct@meta.data %>%
  group_by(integrated_snn_res.0.6) %>%
  tally() %>%
  dplyr::rename('cluster' = integrated_snn_res.0.6)

p2 <- table_clusters_by_samples %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'cluster') %>%
  mutate(cluster = factor(cluster, levels = sort(unique(pacemaker.combined.sct@meta.data$integrated_snn_res.0.6)))) %>%
  ggplot(aes(cluster, value)) +
  geom_bar(aes(fill = variable), position = 'fill', stat = 'identity') +
  geom_text(
    data = temp_labels, aes(x = cluster, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Sample', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Percentage [%]', labels = scales::percent_format(), expand = c(0.01,0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'right',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 10, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/table_clusters_by_samples_cell_percent.pdf", p2, width = 10, height = 5)
```


```{r}
require(rBCS)
pacemaker.combined.sct@meta.data <- select(pacemaker.combined.sct@meta.data, -seurat_clusters)
pacemaker.combined.sct@meta.data <- mutate(pacemaker.combined.sct@meta.data, seurat_clusters = pacemaker.combined.sct@meta.data$integrated_snn_res.0.6)
pacemaker.combined.sct@meta.data <- select(pacemaker.combined.sct@meta.data, Sample = orig.ident, Tissue, seurat_clusters, everything())
pacemaker.combined.sct@meta.data$Tissue <- str_replace(pacemaker.combined.sct@meta.data$Tissue, "D34_", "")
names(pacemaker.combined.sct@meta.data)[2] = "nCount_RNA"
ExportSeurat(pacemaker.combined.sct, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Clustered.pacemaker.combined.sct.bcs")
```


Silhouette plot
```{r}
distance_matrix <- dist(Embeddings(pacemaker.combined.sct[['pca']])[, 1:40])

sil_score_matrix <- matrix(NA, nrow=dim(pacemaker.combined.sct@meta.data)[1], ncol=23)
for(i in 11:33){
  clusters <- pacemaker.combined.sct@meta.data[,i]
  silhouette <- silhouette(as.numeric(clusters), dist = distance_matrix)
  sil_score_matrix[,i-10] <- silhouette[,3]
}


colnames(sil_score_matrix) = paste0("silhouette_score_res.", c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

pacemaker.combined.sct@meta.data <- cbind(pacemaker.combined.sct@meta.data, sil_score_matrix)

pacemaker.combined.sct@meta.data <- select(pacemaker.combined.sct@meta.data, -35)


#clusters <- pacemaker.combined.sct@meta.data$integrated_snn_res.0.6

#save(distance_matrix, file="/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/distance_matrix_for_silhouette.RData")

#X <- bigmemory::as.big.matrix(as.matrix(distance_matrix))
#silhouette <- silhouette(as.numeric(clusters), dist = distance_matrix)

#silhouette <- silhouette(as.numeric(clusters), dist = X)

#pacemaker.combined.sct@meta.data$silhouette_score_res.0.6 <- silhouette[,3]

for(i in 35:57){
#for(i in 35){
  mean_silhouette_score <- mean(pacemaker.combined.sct@meta.data[,i])
  tmp <- select(pacemaker.combined.sct@meta.data, c(i-24, i))
  colnames(tmp) <- c("Cluster", "Sil_score")
  tmp <- mutate(tmp, barcode =  rownames(tmp))
  tmp <- arrange(tmp, Cluster, desc(Sil_score))
  tmp$barcode = factor(tmp$barcode, levels = tmp$barcode)
  p <- 
    ggplot(tmp) +
    #geom_col(aes(barcode, pacemaker.combined.sct@meta.data[,i], fill = pacemaker.combined.sct@meta.data)[,i-24], show.legend = TRUE) +
    geom_col(aes(barcode, Sil_score, fill = Cluster), show.legend = TRUE) +
    geom_hline(yintercept = mean_silhouette_score, color = 'red', linetype = 'dashed') +
    #scale_x_discrete(name = 'Cells') +
    xlab("Clusters") +
    ylab("Silhouette Score") +
    #scale_y_continuous(name = 'Silhouette score') +
    scale_fill_manual(values = custom_colors$discrete, name = 'Clusters') +
    theme_bw() +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
    ggsave(paste0("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/PCs/silhouette_plot_", names(pacemaker.combined.sct@meta.data)[i], ".pdf"), p, height = 5, width = 8)
}

tmp_score <- select(pacemaker.combined.sct@meta.data, 35:57)
library(reshape2)
long_tmp_score <- melt(tmp_score)

long_tmp_score$variable = str_replace(long_tmp_score$variable, "silhouette_score_res.", "")

library(ggpubr)
ggviolin(long_tmp_score, "variable", "value", fill = "variable",
   palette = custom_colors$discrete,
   add = "boxplot", add.params = list(fill = "white"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/PCs/silhouette_violin_plot.pdf", width = 10, height = 5)

#mean_silhouette_score <- mean(pacemaker.combined.sct@meta.data$silhouette_score_res.0.6)

#p <- pacemaker.combined.sct@meta.data %>%
#  mutate(barcode = rownames(.)) %>%
#  arrange(integrated_snn_res.0.6,-silhouette_score_res.0.6) %>%
#  mutate(barcode = factor(barcode, levels = barcode)) %>%
#  ggplot() +
#  geom_col(aes(barcode, silhouette_score_res.0.6, fill = integrated_snn_res.0.6), show.legend = TRUE) +
#  geom_hline(yintercept = mean_silhouette_score, color = 'red', linetype = 'dashed') +
  #scale_x_discrete(name = 'Cells') +
#  xlab("Clusters") +
#  scale_y_continuous(name = 'Silhouette score') +
#  scale_fill_manual(values = custom_colors$discrete, name = 'Clusters') +
#  theme_bw() +
#  theme(
#    axis.title.x = element_blank(),
#    axis.text.x = element_blank(),
#    axis.ticks.x = element_blank(),
#    panel.grid.major = element_blank(),
#    panel.grid.minor = element_blank()
#  )
#ggsave('/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/PCs/silhouette_plot_res.0.6.png', p, height = 5, width = 8)
#ggsave('/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/PCs/silhouette_plot_res.0.6.png', p)

```

Find markers at Res 0.6
```{r}
# find markers for every cluster compared to all remaining cells, report only the positive
# ones
pacemaker.combined.sct.markers <- FindAllMarkers(pacemaker.combined.sct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
save(pacemaker.combined.sct.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.6.RData")

write.table(as.data.frame(pacemaker.combined.sct.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.6.txt", quote = F, sep = "\t")

### This version of results were later move to "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/". The original folder "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/" was used for lenient criteria results. Jun 13 2022.
load("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.6.RData")

ls()
#[1] "filtered.samples.markers"

filtered.pacemaker.samples.markers.fdr <- filter(pacemaker.combined.sct.markers, p_val_adj < 0.01)

dim(filtered.pacemaker.samples.markers.fdr)
#[1] 7587    7
dim(pacemaker.combined.sct.markers)
#[1] 8811    7

filtered.pacemaker.samples.markers.fdr %>%
    group_by(cluster) %>%
    top_n(n = 100, wt = avg_log2FC) -> top100

write.table(as.data.frame(top100),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.6_top100.txt", quote = F, sep = "\t")

####
#filtered.pacemaker.samples.markers.fdr$cluster = as.vector(as.numeric(filtered.pacemaker.samples.markers.fdr$cluster))
top100list <- matrix(0, nrow = 100, ncol = 20)
top100_df <- as.data.frame(top100)
for(i in seq(0,19,1)){
  j <- i+1
  top100list[1:length(top100_df[top100_df$cluster == i,7]),j] <- top100_df[top100_df$cluster == i,7]
}
colnames(top100list) <- paste0("C_", seq(0,19,1))

#top100list <- select(as.data.frame(top100list), c("C_18","C_6","C_8","C_3","C_15","C_13","C_21","C_12","C_5","C_2","C_10","C_9","C_16","C_26","C_27","C_19","C_20","C_1","C_24","C_23","C_25","C_17","C_22","C_14","C_4","C_7","C_11","C_0"))

write.table(as.data.frame(top100list),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.6_top100list.txt", quote = F, sep = "\t", row.names = F)
```


Find markers at Res 0.2
```{r}
# find markers for every cluster compared to all remaining cells, report only the positive
# ones
Idents(object = pacemaker.combined.sct) <- "integrated_snn_res.0.2"
pacemaker.combined.sct.markers <- FindAllMarkers(pacemaker.combined.sct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
save(pacemaker.combined.sct.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.2.RData")

write.table(as.data.frame(pacemaker.combined.sct.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.2.txt", quote = F, sep = "\t")

filtered.pacemaker.samples.markers.fdr <- filter(pacemaker.combined.sct.markers, p_val_adj < 0.01)

table(filtered.pacemaker.samples.markers.fdr$cluster)
#   0    1    2    3    4    5    6    7    8    9 
# 198  281  848 1166  373  327  230  469  673  439

filtered.pacemaker.samples.markers.fdr.log2fc <- filter(pacemaker.combined.sct.markers, p_val_adj < 0.01 & avg_log2FC > log2(1.5))

table(filtered.pacemaker.samples.markers.fdr.log2fc$cluster)
#   0    1    2    3    4    5    6    7    8    9 
# 159  242  800 1108  332  289  217  460  657  430 

dim(filtered.pacemaker.samples.markers.fdr.log2fc)
#[1] 4694    7
dim(filtered.pacemaker.samples.markers.fdr)
#[1] 5004    7
dim(pacemaker.combined.sct.markers)
#[1] 5748    7

filtered.pacemaker.samples.markers.fdr.log2fc %>%
    group_by(cluster) %>%
    slice_max(n = 100, order_by = -p_val_adj) -> top100

filtered.pacemaker.samples.markers.fdr.log2fc %>%
    group_by(cluster) %>%
    slice_head(n = 100) -> top100

write.table(as.data.frame(top100),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.2_top100_fdr.txt", quote = F, sep = "\t")
####
#filtered.pacemaker.samples.markers.fdr$cluster = as.vector(as.numeric(filtered.pacemaker.samples.markers.fdr$cluster))
top100list <- matrix(0, nrow = 100, ncol = 10)
top100_df <- as.data.frame(top100)
for(i in seq(0,9,1)){
  j <- i+1
  top100list[1:length(top100_df[top100_df$cluster == i,7]),j] <- top100_df[top100_df$cluster == i,7]
}
colnames(top100list) <- paste0("C_", seq(0,9,1))

#top100list <- select(as.data.frame(top100list), c("C_18","C_6","C_8","C_3","C_15","C_13","C_21","C_12","C_5","C_2","C_10","C_9","C_16","C_26","C_27","C_19","C_20","C_1","C_24","C_23","C_25","C_17","C_22","C_14","C_4","C_7","C_11","C_0"))

write.table(as.data.frame(top100list),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.2_top100list.txt", quote = F, sep = "\t", row.names = F)
```


```{r}
Idents(object = pacemaker.combined.sct) <- "integrated_snn_res.0.2"
DimPlot(pacemaker.combined.sct,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        cols = custom_colors$discrete, 
        label.size = 10)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/UMAP.res.0.2.D34.cells.5000.features.pdf", width = 10, height = 8)

DimPlot(pacemaker.combined.sct,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        cols = custom_colors$discrete, 
        group.by = "integrated_snn_res.0.3",
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/UMAP.res.0.3.D34.cells.5000.features.pdf", width = 10, height = 8)
```


#####################################
Differential expression analysis using assay "RNA" slot "data".
```{r}
## Try to use RNA assay for differential expression analysis.
pacemaker.combined.sct.RNA <- pacemaker.combined.sct
Idents(object = pacemaker.combined.sct.RNA) <- "integrated_snn_res.0.2"
DefaultAssay(pacemaker.combined.sct.RNA) <- "RNA"
pacemaker.combined.sct.RNA <- NormalizeData(pacemaker.combined.sct.RNA, normalization.method = "LogNormalize", scale.factor = 10000)
pacemaker.combined.sct.RNA[["RNA"]]@data

C0_vs_C1_RNA <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "0", ident.2 = "1", min.pct = 0.25, assay = "RNA", slot = "data")
write.table(C0_vs_C1_RNA,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C0_vs_C1_RNA_Assay.txt", quote = F, sep = "\t", row.names = T)
dim(C0_vs_C1_RNA)
#[1] 401   5
C0_vs_C1_Filter <- filter(C0_vs_C1_RNA, abs(avg_log2FC) > 0.25 & p_val_adj < 0.05)
dim(C0_vs_C1_Filter)
#[1] 400   5
dim(filter(C0_vs_C1_Filter, avg_log2FC > 0))

C0_vs_C2_RNA <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "0", ident.2 = "2", min.pct = 0.25, assay = "RNA", slot = "data")
write.table(C0_vs_C2_RNA,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C0_vs_C2_RNA_Assay.txt", quote = F, sep = "\t", row.names = T)
dim(C0_vs_C2_RNA)
#[1] 962   5
C0_vs_C2_Filter <- filter(C0_vs_C2_RNA, abs(avg_log2FC) > 0.25 & p_val_adj < 0.05)
dim(C0_vs_C2_Filter)
#[1] 962   5
dim(filter(C0_vs_C2_Filter, avg_log2FC > 0))

C2_vs_C1_RNA <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "2", ident.2 = "1", min.pct = 0.25, assay = "RNA", slot = "data")
write.table(C2_vs_C1_RNA,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C2_vs_C1_RNA_Assay.txt", quote = F, sep = "\t", row.names = T)
dim(C2_vs_C1_RNA)
#[1] 1468    5
C2_vs_C1_Filter <- filter(C2_vs_C1_RNA, abs(avg_log2FC) > 0.25 & p_val_adj < 0.05)
dim(C2_vs_C1_Filter)
#[1] 1466    5
dim(filter(C2_vs_C1_Filter, avg_log2FC > 0))

#######################################################################################################################################################3
##### Try logistic regression for differential expression.
C0_vs_C1_RNA_LR <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "0", ident.2 = "1", min.pct = 0.25, assay = "RNA", slot = "data", test.use = "LR", latent.vars = "orig.ident")
write.table(C0_vs_C1_RNA_LR,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C0_vs_C1_RNA_Assay_LR.txt", quote = F, sep = "\t", row.names = T)
##### The logFCs are the same between these two tests. Only the p-values are different, but all 401 genes has p-values < 0.05. So decided to use default wilcox test.

##### Try MAST for differential expression.
C0_vs_C1_RNA_MAST <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "0", ident.2 = "1", min.pct = 0.25, assay = "RNA", slot = "data", test.use = "MAST", latent.vars = "orig.ident")
write.table(C0_vs_C1_RNA_MAST,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C0_vs_C1_RNA_Assay_MAST.txt", quote = F, sep = "\t", row.names = T)

##### Try DESeq2 for differential expression. Takes very long to run.
C0_vs_C1_RNA_DESeq2 <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "0", ident.2 = "1", min.pct = 0.25, assay = "RNA", slot = "data", test.use = "DESeq2")
write.table(C0_vs_C1_RNA_DESeq2,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C0_vs_C1_RNA_Assay_DESeq2.txt", quote = F, sep = "\t", row.names = T)

##### Try Students t-test for differential expression.
C0_vs_C1_RNA_ttest <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "0", ident.2 = "1", min.pct = 0.25, assay = "RNA", slot = "data", test.use = "t")
write.table(C0_vs_C1_RNA_ttest,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C0_vs_C1_RNA_Assay_ttest.txt", quote = F, sep = "\t", row.names = T)

##### Try poisson for differential expression.
C0_vs_C1_RNA_poisson <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "0", ident.2 = "1", min.pct = 0.25, assay = "RNA", slot = "data", test.use = "poisson", latent.vars = "orig.ident")
write.table(C0_vs_C1_RNA_poisson,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C0_vs_C1_RNA_Assay_poisson.txt", quote = F, sep = "\t", row.names = T)

##### Try negbinom for differential expression. Takes longer time to run.
C0_vs_C1_RNA_negbinom <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "0", ident.2 = "1", min.pct = 0.25, assay = "RNA", slot = "data", test.use = "negbinom", latent.vars = "orig.ident")
write.table(C0_vs_C1_RNA_negbinom,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C0_vs_C1_RNA_Assay_negbinom.txt", quote = F, sep = "\t", row.names = T)
###### negbinom, poisson and DESeq2 report more genes than others, but skewed in the number of upregulated and downregulated genes.
#######################################################################################################################################################3
```


Plot genes from differential expression testing between clusters.
```{r}
##### C0_vs_C1
C0_vs_C1_Filter <- mutate(C0_vs_C1_Filter, Gene = rownames(C0_vs_C1_Filter))

# Single cell heatmap of feature expression
pacemaker.combined.sct.RNA.subset <- subset(pacemaker.combined.sct.RNA, subset = integrated_snn_res.0.2 == "0" | integrated_snn_res.0.2 == "1")
mat<- pacemaker.combined.sct.RNA.subset[["RNA"]]@data[rownames(C0_vs_C1_Filter),] %>% as.matrix()
#pacemaker.combined.sct.RNA.subset[["SCT"]]@data[unique(rownames(combine_DEGs)),]
## scale the rows
mat<- t(scale(t(mat)))
cluster_anno<- pacemaker.combined.sct.RNA.subset@meta.data$integrated_snn_res.0.2
### Annotate certain genes.
ha = rowAnnotation(foo = anno_mark(at = c(2,9,29,35,53,62,65,100,89,109,138,141,165,214,225), labels = rownames(C0_vs_C1_Filter)[c(2,9,29,35,53,62,65,100,89,109,138,141,165,214,225)]))

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_C1_DEGs_heatmap.pdf", height = 8, width = 10)
ht <- Heatmap(mat, name = "Expression",  
        column_split = factor(cluster_anno),
        cluster_columns = FALSE,
        show_column_dend = FALSE,
        cluster_column_slices = TRUE,
        column_title_gp = gpar(fontsize = 8),
        column_gap = unit(0, "mm"),
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        #col = col_fun,
        row_names_gp = gpar(fontsize = 4),
        column_title_rot = 0,
        #top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = scales::hue_pal()(7)))),
        #HeatmapAnnotation(bar = cluster_anno, col = list(bar = c("ACM" = "red", "VCM" = "green"))),
        top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = paletteer_d("ggsci::planetexpress_futurama")[c(1,2)]))),
        #right_annotation = HeatmapAnnotation(foo = anno_barplot(ACM_vs_VCM_DE_order$avg_log2FC, baseline = 0)),
        show_column_names = FALSE,
        show_row_names = FALSE,
        use_raster = TRUE,
        raster_quality = 4,
        right_annotation = ha,
        col = colorRamp2(c(-4, 0, 4), c("#56B4E9", "#F7FAFF", "#E69F00")))
draw(ht)
dev.off()


##### C0_vs_C2
C0_vs_C2_Filter <- mutate(C0_vs_C2_Filter, Gene = rownames(C0_vs_C2_Filter))
C0_vs_C2_Filter_order <- arrange(C0_vs_C2_Filter, avg_log2FC)
# Single cell heatmap of feature expression
pacemaker.combined.sct.RNA.subset <- subset(pacemaker.combined.sct.RNA, subset = integrated_snn_res.0.2 == "0" | integrated_snn_res.0.2 == "2")
mat<- pacemaker.combined.sct.RNA.subset[["RNA"]]@data[rownames(C0_vs_C2_Filter_order),] %>% as.matrix()
#pacemaker.combined.sct.RNA.subset[["SCT"]]@data[unique(rownames(combine_DEGs)),]
## scale the rows
mat<- t(scale(t(mat)))
cluster_anno<- pacemaker.combined.sct.RNA.subset@meta.data$integrated_snn_res.0.2
### Annotate certain genes.
ha = rowAnnotation(foo = anno_mark(at = c(1:20,93, 699,782,895,915,(dim(C0_vs_C2_Filter_order)[1]-19):dim(C0_vs_C2_Filter_order)[1]), labels = rownames(C0_vs_C2_Filter_order)[c(1:20,93, 699,782,895,915,(dim(C0_vs_C2_Filter_order)[1]-19):dim(C0_vs_C2_Filter_order)[1])]))

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_C2_DEGs_heatmap.pdf", height = 8, width = 10)
ht <- Heatmap(mat, name = "Expression",  
        column_split = factor(cluster_anno),
        cluster_columns = FALSE,
        show_column_dend = FALSE,
        cluster_column_slices = TRUE,
        column_title_gp = gpar(fontsize = 8),
        column_gap = unit(0, "mm"),
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        #col = col_fun,
        row_names_gp = gpar(fontsize = 4),
        column_title_rot = 0,
        #top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = scales::hue_pal()(7)))),
        #HeatmapAnnotation(bar = cluster_anno, col = list(bar = c("ACM" = "red", "VCM" = "green"))),
        top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = paletteer_d("ggsci::planetexpress_futurama")[c(1,3)]))),
        #right_annotation = HeatmapAnnotation(foo = anno_barplot(ACM_vs_VCM_DE_order$avg_log2FC, baseline = 0)),
        show_column_names = FALSE,
        show_row_names = FALSE,
        use_raster = TRUE,
        raster_quality = 4,
        right_annotation = ha,
        col = colorRamp2(c(-4, 0, 4), c("#006A8E", "#F7FAFF", "#B1283A")))
draw(ht)
dev.off()

##### C2_vs_C1
C2_vs_C1_Filter <- mutate(C2_vs_C1_Filter, Gene = rownames(C2_vs_C1_Filter))
C2_vs_C1_Filter_order <- arrange(C2_vs_C1_Filter, avg_log2FC)
# Single cell heatmap of feature expression
pacemaker.combined.sct.RNA.subset <- subset(pacemaker.combined.sct.RNA, subset = integrated_snn_res.0.2 == "1" | integrated_snn_res.0.2 == "2")
mat<- pacemaker.combined.sct.RNA.subset[["RNA"]]@data[rownames(C2_vs_C1_Filter_order),] %>% as.matrix()
#pacemaker.combined.sct.RNA.subset[["SCT"]]@data[unique(rownames(combine_DEGs)),]
## scale the rows
mat<- t(scale(t(mat)))
cluster_anno<- pacemaker.combined.sct.RNA.subset@meta.data$integrated_snn_res.0.2
### Annotate certain genes.
ha = rowAnnotation(foo = anno_mark(at = c(1:20,(dim(C2_vs_C1_Filter_order)[1]-19):dim(C2_vs_C1_Filter_order)[1]), labels = rownames(C2_vs_C1_Filter_order)[c(1:20,(dim(C2_vs_C1_Filter_order)[1]-19):dim(C2_vs_C1_Filter_order)[1])]))

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C2_C1_DEGs_heatmap.pdf", height = 8, width = 10)
ht <- Heatmap(mat, name = "Expression",  
        column_split = factor(cluster_anno),
        cluster_columns = FALSE,
        show_column_dend = FALSE,
        cluster_column_slices = TRUE,
        column_title_gp = gpar(fontsize = 8),
        column_gap = unit(0, "mm"),
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        #col = col_fun,
        row_names_gp = gpar(fontsize = 4),
        column_title_rot = 0,
        #top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = scales::hue_pal()(7)))),
        #HeatmapAnnotation(bar = cluster_anno, col = list(bar = c("ACM" = "red", "VCM" = "green"))),
        top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = paletteer_d("ggsci::planetexpress_futurama")[c(2:3)]))),
        #right_annotation = HeatmapAnnotation(foo = anno_barplot(ACM_vs_VCM_DE_order$avg_log2FC, baseline = 0)),
        show_column_names = FALSE,
        show_row_names = FALSE,
        use_raster = TRUE,
        raster_quality = 4,
        right_annotation = ha,
        col = colorRamp2(c(-4, 0, 4), c("#052F61", "#F7FAFF", "#A50E82")))
draw(ht)
dev.off()

#### All DEGs together. Not looking good.
C0_vs_C1_Filter <- mutate(C0_vs_C1_Filter, Gene = rownames(C0_vs_C1_Filter))
C0_vs_C2_Filter <- mutate(C0_vs_C2_Filter, Gene = rownames(C0_vs_C2_Filter))
C2_vs_C1_Filter <- mutate(C2_vs_C1_Filter, Gene = rownames(C2_vs_C1_Filter))
combine_DEGs <- rbind(C0_vs_C1_Filter, C0_vs_C2_Filter, C2_vs_C1_Filter)
write.table(unique(combine_DEGs$Gene),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/unique_combine_DEGs.txt", quote = F, sep = "\t", row.names = F, col.names = F)
combine_DEGs <- unique(combine_DEGs$Gene)

# Single cell heatmap of feature expression
pacemaker.combined.sct.RNA.subset <- subset(pacemaker.combined.sct.RNA, subset = integrated_snn_res.0.2 == "0" | integrated_snn_res.0.2 == "1" | integrated_snn_res.0.2 == "2")
mat<- pacemaker.combined.sct.RNA.subset[["RNA"]]@data[combine_DEGs,] %>% as.matrix()
#pacemaker.combined.sct.RNA.subset[["SCT"]]@data[unique(rownames(combine_DEGs)),]
## scale the rows
mat<- t(scale(t(mat)))
cluster_anno<- pacemaker.combined.sct.RNA.subset@meta.data$integrated_snn_res.0.2
### Annotate certain genes.
ha = rowAnnotation(foo = anno_mark(at = c(2,9,14,29,35,53,62,65,89,100,109,118,138,141,165,214,225,447,452,458,508,550,665,1103), labels = combine_DEGs[c(2,9,14,29,35,53,62,65,89,100,109,118,138,141,165,214,225,447,452,458,508,550,665,1103)]))
#ha = rowAnnotation(foo = anno_mark(at = c(1, 4, 11, 31, 49, 131, 167, 181, 191, 217, 222), labels = c("NPPB", "NPPA", "MYL3", "MYL7", "MYL6", "TBX5", "TBX3", "MYH11", "FOS", "TBX18", "SHOX2")), bar = anno_barplot(ACM_vs_VCM_DE_order$avg_log2FC))
pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_C1_C2_DEGs_heatmap.pdf", height = 10, width = 10)
ht <- Heatmap(mat, name = "Expression",  
        column_split = factor(cluster_anno),
        cluster_columns = FALSE,
        show_column_dend = FALSE,
        cluster_column_slices = TRUE,
        column_title_gp = gpar(fontsize = 8),
        column_gap = unit(0, "mm"),
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        #col = col_fun,
        row_names_gp = gpar(fontsize = 4),
        column_title_rot = 0,
        #top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = scales::hue_pal()(7)))),
        #HeatmapAnnotation(bar = cluster_anno, col = list(bar = c("ACM" = "red", "VCM" = "green"))),
        top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = paletteer_d("ggsci::planetexpress_futurama")))),
        #right_annotation = HeatmapAnnotation(foo = anno_barplot(ACM_vs_VCM_DE_order$avg_log2FC, baseline = 0)),
        show_column_names = FALSE,
        show_row_names = FALSE,
        use_raster = TRUE,
        raster_quality = 4,
        right_annotation = ha,
        col = colorRamp2(c(-4, 0, 4), c("#56B4E9", "#F7FAFF", "#E69F00")))
draw(ht)
dev.off()
```


```{r}
# Plot top 25 pathways
pathways <- read.table("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DEGs_Metascape.txt", sep = "\t", quote = "", header = T, comment.char = "", check.names = F)

## Plot FDR values rather than p-values for GO terms.
pathways_fdr <- as.data.frame(apply(pathways[1:6], 2, function(x){log10(p.adjust(10^x))}))
pathways_fdr <- mutate(pathways_fdr, Category = pathways$Category, Description = pathways$Description)
GO_fdr <- filter(pathways_fdr, Category == "GO Biological Processes")

### C0_vs_C1
C0_vs_C1_GO_Up_fdr <- GO_fdr %>% arrange(C0_vs_C1_Up) %>% head(n=10)
C0_vs_C1_GO_Down_fdr <- GO_fdr %>% arrange(C0_vs_C1_Down) %>% head(n=10)
C0_vs_C1_Up_Down_top10_fdr <- GO_fdr %>% filter(Description %in% unique(c(C0_vs_C1_GO_Up_fdr$Description, C0_vs_C1_GO_Down_fdr$Description))) %>% select(C0_vs_C1_Down, C0_vs_C1_Up, Description) %>% arrange(desc(C0_vs_C1_Up))
C0_vs_C1_Up_Down_top10_long_fdr <- melt(C0_vs_C1_Up_Down_top10_fdr)
C0_vs_C1_Up_Down_top10_long_fdr$variable <- factor(C0_vs_C1_Up_Down_top10_long_fdr$variable, levels = c("C0_vs_C1_Up","C0_vs_C1_Down"))
C0_vs_C1_Up_Down_top10_long_fdr$Description <- factor(C0_vs_C1_Up_Down_top10_long_fdr$Description, levels = unique(C0_vs_C1_Up_Down_top10_fdr$Description))
ggplot(C0_vs_C1_Up_Down_top10_long_fdr,aes(x=variable, y=Description)) + geom_point(aes(size=-value), colour = "#6BA3D6") + xlab("DEG Sets") + ylab("GO Pathways") + labs(size="-LogFDR") + theme(axis.text=  element_text(color = "black"), axis.title =  element_text(size = 9))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C1_Up_Down_Top_10_Pathways_fdr.pdf", width = 6, height = 8)

# Only plot top 10 pathways in up DEGs.
C0_vs_C1_GO_Up_fdr$Description = factor(C0_vs_C1_GO_Up_fdr$Description, levels = rev(C0_vs_C1_GO_Up_fdr$Description))
ggplot(C0_vs_C1_GO_Up_fdr, aes(x = -C0_vs_C1_Up, y=Description)) + geom_bar(stat="identity", fill="#A39FC9") + xlab("-log10(FDR)") + ylab("GO Biological Processes") + theme_classic()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C1_Up_Top_10_Pathways_fdr.pdf", width = 6, height = 6)
# Only plot top 10 pathways in Down DEGs.
C0_vs_C1_GO_Down_fdr$Description = factor(C0_vs_C1_GO_Down_fdr$Description, levels = rev(C0_vs_C1_GO_Down_fdr$Description))
ggplot(C0_vs_C1_GO_Down_fdr, aes(x = -C0_vs_C1_Down, y=Description)) + geom_bar(stat="identity", fill="#92AA4C") + xlab("-log10(FDR)") + ylab("GO Biological Processes") + theme_classic()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C1_Down_Top_10_Pathways_fdr.pdf", width = 6, height = 6)


### C0_vs_C2
C0_vs_C2_GO_Up_fdr <- GO_fdr %>% arrange(C0_vs_C2_Up) %>% head(n=10)
C0_vs_C2_GO_Down_fdr <- GO_fdr %>% arrange(C0_vs_C2_Down) %>% head(n=10)
C0_vs_C2_Up_Down_top10_fdr <- GO_fdr %>% filter(Description %in% unique(c(C0_vs_C2_GO_Up_fdr$Description, C0_vs_C2_GO_Down_fdr$Description))) %>% select(C0_vs_C2_Down, C0_vs_C2_Up, Description) %>% arrange(desc(C0_vs_C2_Up))
C0_vs_C2_Up_Down_top10_long_fdr <- melt(C0_vs_C2_Up_Down_top10_fdr)
C0_vs_C2_Up_Down_top10_long_fdr$variable <- factor(C0_vs_C2_Up_Down_top10_long_fdr$variable, levels = c("C0_vs_C2_Up","C0_vs_C2_Down"))
C0_vs_C2_Up_Down_top10_long_fdr$Description <- factor(C0_vs_C2_Up_Down_top10_long_fdr$Description, levels = unique(C0_vs_C2_Up_Down_top10_fdr$Description))
ggplot(C0_vs_C2_Up_Down_top10_long_fdr,aes(x=variable, y=Description)) + geom_point(aes(size=-value), colour = "#6BA3D6") + xlab("DEG Sets") + ylab("GO Pathways") + labs(size="-LogFDR") + theme(axis.text=  element_text(color = "black"), axis.title =  element_text(size = 9))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C2_Up_Down_Top_10_Pathways_fdr.pdf", width = 6, height = 8)

# Only plot top 10 pathways in up DEGs.
C0_vs_C2_GO_Up_fdr$Description = factor(C0_vs_C2_GO_Up_fdr$Description, levels = rev(C0_vs_C2_GO_Up_fdr$Description))
ggplot(C0_vs_C2_GO_Up_fdr, aes(x = -C0_vs_C2_Up, y=Description)) + geom_bar(stat="identity", fill="#A39FC9") + xlab("-log10(FDR)") + ylab("GO Biological Processes") + theme_classic()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C2_Up_Top_10_Pathways_fdr.pdf", width = 6, height = 6)
# Only plot top 10 pathways in Down DEGs.
C0_vs_C2_GO_Down_fdr$Description = factor(C0_vs_C2_GO_Down_fdr$Description, levels = rev(C0_vs_C2_GO_Down_fdr$Description))
ggplot(C0_vs_C2_GO_Down_fdr, aes(x = -C0_vs_C2_Down, y=Description)) + geom_bar(stat="identity", fill="#92AA4C") + xlab("-log10(FDR)") + ylab("GO Biological Processes") + theme_classic()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C2_Down_Top_10_Pathways_fdr.pdf", width = 6, height = 6)


### C2_vs_C1
C2_vs_C1_GO_Up_fdr <- GO_fdr %>% arrange(C2_vs_C1_Up) %>% head(n=10)
C2_vs_C1_GO_Down_fdr <- GO_fdr %>% arrange(C2_vs_C1_Down) %>% head(n=10)
C2_vs_C1_Up_Down_top10_fdr <- GO_fdr %>% filter(Description %in% unique(c(C2_vs_C1_GO_Up_fdr$Description, C2_vs_C1_GO_Down_fdr$Description))) %>% select(C2_vs_C1_Down, C2_vs_C1_Up, Description) %>% arrange(desc(C2_vs_C1_Up))
C2_vs_C1_Up_Down_top10_long_fdr <- melt(C2_vs_C1_Up_Down_top10_fdr)
C2_vs_C1_Up_Down_top10_long_fdr$variable <- factor(C2_vs_C1_Up_Down_top10_long_fdr$variable, levels = c("C2_vs_C1_Up","C2_vs_C1_Down"))
C2_vs_C1_Up_Down_top10_long_fdr$Description <- factor(C2_vs_C1_Up_Down_top10_long_fdr$Description, levels = unique(C2_vs_C1_Up_Down_top10_fdr$Description))
ggplot(C2_vs_C1_Up_Down_top10_long_fdr,aes(x=variable, y=Description)) + geom_point(aes(size=-value), colour = "#6BA3D6") + xlab("DEG Sets") + ylab("GO Pathways") + labs(size="-LogFDR") + theme(axis.text=  element_text(color = "black"), axis.title =  element_text(size = 9))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C2_vs_C1_Up_Down_Top_10_Pathways_fdr.pdf", width = 6, height = 8)

# Only plot top 10 pathways in up DEGs.
C2_vs_C1_GO_Up_fdr$Description = factor(C2_vs_C1_GO_Up_fdr$Description, levels = rev(C2_vs_C1_GO_Up_fdr$Description))
ggplot(C2_vs_C1_GO_Up_fdr, aes(x = -C2_vs_C1_Up, y=Description)) + geom_bar(stat="identity", fill="#A39FC9") + xlab("-log10(FDR)") + ylab("GO Biological Processes") + theme_classic()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C2_vs_C1_Up_Top_10_Pathways_fdr.pdf", width = 6, height = 6)
# Only plot top 10 pathways in Down DEGs.
C2_vs_C1_GO_Down_fdr$Description = factor(C2_vs_C1_GO_Down_fdr$Description, levels = rev(C2_vs_C1_GO_Down_fdr$Description))
ggplot(C2_vs_C1_GO_Down_fdr, aes(x = -C2_vs_C1_Down, y=Description)) + geom_bar(stat="identity", fill="#92AA4C") + xlab("-log10(FDR)") + ylab("GO Biological Processes") + theme_classic()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C2_vs_C1_Down_Top_10_Pathways_fdr.pdf", width = 6, height = 6)
```


Find markers at Res 0.3 using RNA assay.
```{r}
## Try to use RNA assay for differential expression analysis.
#pacemaker.combined.sct <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Clustered.pacemaker.combined.sct.rds")
pacemaker.combined.sct.RNA <- pacemaker.combined.sct
Idents(object = pacemaker.combined.sct.RNA) <- "integrated_snn_res.0.3"
DefaultAssay(pacemaker.combined.sct.RNA) <- "RNA"
pacemaker.combined.sct.RNA <- NormalizeData(pacemaker.combined.sct.RNA, normalization.method = "LogNormalize", scale.factor = 10000)
pacemaker.combined.sct.RNA[["RNA"]]@data

pacemaker.combined.sct.RNA.markers <- FindAllMarkers(pacemaker.combined.sct.RNA, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, assay = "RNA", slot = "data")

save(pacemaker.combined.sct.RNA.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.3_RNA_Assay.RData")

write.table(as.data.frame(pacemaker.combined.sct.RNA.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.3_RNA_Assay.txt", quote = F, sep = "\t")

pacemaker.combined.sct.RNA.markers.fdr <- filter(pacemaker.combined.sct.RNA.markers, p_val_adj < 0.01)

table(pacemaker.combined.sct.RNA.markers.fdr$cluster)
#  0   1   2   3   4   5   6   7   8   9  10  11  12  13 
#163 170  68 603 316 290 538 711 659 754 223 982 738 579 

pacemaker.combined.sct.RNA.markers.fdr %>%
    group_by(cluster) %>%
    slice_head(n = 200) -> top200

write.table(as.data.frame(top200),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.3.RNA_top200.txt", quote = F, sep = "\t")

top200list <- matrix(0, nrow = 200, ncol = 14)
top200_df <- as.data.frame(top200)
for(i in seq(0,13,1)){
  j <- i+1
  top200list[1:length(top200_df[top200_df$cluster == i,7]),j] <- top200_df[top200_df$cluster == i,7]
}
colnames(top200list) <- paste0("C_", seq(0,13,1))

write.table(as.data.frame(top200list),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.3.RNA_top200list.txt", quote = F, sep = "\t", row.names = F)


### Plot top 5 markers for each cluster.
pacemaker.combined.sct.RNA.markers.fdr %>%
    group_by(cluster) %>%
    slice_head(n = 5) -> top5

DotPlot_scCustom(pacemaker.combined.sct.RNA, features = unique(top5$gene), flip_axes = T, remove_axis_titles = FALSE, colors_use = rev(viridis(1000))) + xlab("Genes") + ylab("Clusters")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Top5.markers_dotplot_Res.0.3.RNA_clusters_RNA_Assay.pdf", width = 6, height = 10)
```


```{r}
C1_vs_C2_RNA_res0.3 <- FindMarkers(pacemaker.combined.sct.RNA, ident.1 = "1", ident.2 = "2", min.pct = 0.25, assay = "RNA", slot = "data")

write.table(C1_vs_C2_RNA_res0.3,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DE_C1_vs_C2_RNA_Assay_Res0.3.txt", quote = F, sep = "\t", row.names = T)
dim(C1_vs_C2_RNA_res0.3)
#[1] 373   5
C1_vs_C2_RNA_res0.3_Filter <- filter(C1_vs_C2_RNA_res0.3, abs(avg_log2FC) > 0.25 & p_val_adj < 0.05)
dim(C1_vs_C2_RNA_res0.3_Filter)
#[1] 372   5
dim(filter(C1_vs_C2_RNA_res0.3_Filter, avg_log2FC > 0))
#[1] 263   5
```


Plot genes from differential expression testing between clusters.
```{r}
C1_vs_C2_RNA_res0.3_Filter <- mutate(C1_vs_C2_RNA_res0.3_Filter, Gene = rownames(C1_vs_C2_RNA_res0.3_Filter))

# Single cell heatmap of feature expression
C1_vs_C2.RNA.subset <- subset(pacemaker.combined.sct.RNA, subset = integrated_snn_res.0.3 == "1" | integrated_snn_res.0.3 == "2")
mat<- C1_vs_C2.RNA.subset[["RNA"]]@data[rownames(C1_vs_C2_RNA_res0.3_Filter),] %>% as.matrix()
#tail.head.RNA.subset[["SCT"]]@data[unique(rownames(combine_DEGs)),]
## scale the rows
mat<- t(scale(t(mat)))
cluster_anno<- C1_vs_C2.RNA.subset@meta.data$integrated_snn_res.0.3
### Annotate certain genes.
ha = rowAnnotation(foo = anno_mark(at = c(19,20,34,38,40,43,46,50,53,103,106,143,150,177,188,200,206,211,244,265,267,273,282,314,319,326,342,370,372), labels = rownames(C1_vs_C2_RNA_res0.3_Filter)[c(19,20,34,38,40,43,46,50,53,103,106,143,150,177,188,200,206,211,244,265,267,273,282,314,319,326,342,370,372)]))

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/tail.head_DEGs_heatmap.pdf", height = 8, width = 10)
ht <- Heatmap(mat, name = "Expression",  
        column_split = factor(cluster_anno),
        cluster_columns = FALSE,
        show_column_dend = FALSE,
        cluster_column_slices = TRUE,
        #column_title_gp = gpar(fontsize = 8),
        column_gap = unit(0, "mm"),
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        #col = col_fun,
        row_names_gp = grid::gpar(fontsize = 0),
        column_title_rot = 0,
        #top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = scales::hue_pal()(7)))),
        #HeatmapAnnotation(bar = cluster_anno, col = list(bar = c("ACM" = "red", "VCM" = "green"))),
        top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = paletteer_d("ggsci::planetexpress_futurama")[c(1,2,3,4)]))),
        #right_annotation = HeatmapAnnotation(foo = anno_barplot(ACM_vs_VCM_DE_order$avg_log2FC, baseline = 0)),
        show_column_names = FALSE,
        show_row_names = FALSE,
        use_raster = TRUE,
        raster_quality = 4,
        right_annotation = ha,
        col = colorRamp2(c(-4, 0, 4), c("#56B4E9", "#F7FAFF", "#E69F00")))
draw(ht)
dev.off()
```

Remove cluster 4 (high mitochondrial expression), 6 (low gene count), 8 (low gene count), 9 (stressed cardiomyocytes), 10 (unknown) , 12 (unknown) at Res 0.3.
```{r}
pacemaker.combined.sct <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Clustered.pacemaker.combined.sct.rds")
Idents(object = pacemaker.combined.sct) <- "integrated_snn_res.0.3"

pacemaker.combined.sct.subset <- subset(pacemaker.combined.sct, subset = integrated_snn_res.0.3 == "0" | integrated_snn_res.0.3 == "1" | integrated_snn_res.0.3 == "2" | integrated_snn_res.0.3 == "3" | integrated_snn_res.0.3 == "5" | integrated_snn_res.0.3 == "7" | integrated_snn_res.0.3 == "11"  | integrated_snn_res.0.3 == "13")

table(pacemaker.combined.sct.subset$integrated_snn_res.0.3)
pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset =  as.vector(pacemaker.combined.sct.subset$integrated_snn_res.0.3)
pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset=="5"]=4
pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset=="7"]=5
pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset=="11"]=6
pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset=="13"]=7
table(pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset)

pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset = factor(pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset, levels = 0:7)
table(pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset)
Idents(object = pacemaker.combined.sct.subset) <- "integrated_snn_res.0.3.subset"

saveRDS(pacemaker.combined.sct.subset, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Filtered.Clustered.pacemaker.combined.sct.rds")
#pacemaker.combined.sct.subset <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Filtered.Clustered.pacemaker.combined.sct.rds")

pacemaker.combined.sct.subset$Annotation =  as.vector(pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset)
pacemaker.combined.sct.subset$Annotation[pacemaker.combined.sct.subset$Annotation=="0"]="ACM"
pacemaker.combined.sct.subset$Annotation[pacemaker.combined.sct.subset$Annotation=="1"]="SAN_Head"
pacemaker.combined.sct.subset$Annotation[pacemaker.combined.sct.subset$Annotation=="2"]="SAN_Tail"
pacemaker.combined.sct.subset$Annotation[pacemaker.combined.sct.subset$Annotation=="3"]="Precursor"
pacemaker.combined.sct.subset$Annotation[pacemaker.combined.sct.subset$Annotation=="4"]="SAN_Tran_1"
pacemaker.combined.sct.subset$Annotation[pacemaker.combined.sct.subset$Annotation=="5"]="SAN_Tran_2"
pacemaker.combined.sct.subset$Annotation[pacemaker.combined.sct.subset$Annotation=="6"]="Proepicardial"
pacemaker.combined.sct.subset$Annotation[pacemaker.combined.sct.subset$Annotation=="7"]="Visceral_Neuron"

pacemaker.combined.sct.subset <- subset(pacemaker.combined.sct.subset, subset = Annotation != "Visceral_Neuron")
table(pacemaker.combined.sct.subset$Annotation)
pacemaker.combined.sct.subset$Annotation = factor(pacemaker.combined.sct.subset$Annotation, levels = unique(pacemaker.combined.sct.subset$Annotation))
Idents(object = pacemaker.combined.sct.subset) <- "Annotation"
table(pacemaker.combined.sct.subset$Annotation,pacemaker.combined.sct.subset$integrated_snn_res.0.3.subset)

DimPlot(pacemaker.combined.sct.subset,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        #cols = custom_colors$discrete, 
        cols = paletteer_d("ggthemes::Tableau_10"), 
        pt.size = 1,
        label.size = 10)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/UMAP.res.0.3.subset.pdf", width = 10, height = 8)

# Violin plot
VlnPlot(pacemaker.combined.sct.subset, c("TBX18", "SHOX2", "BMP4", "ISL1", "TBX3", "NKX2-5"), group.by = "Annotation", pt.size = 0, cols = custom_colors$discrete, ncol = 2)
#VlnPlot(pacemaker.combined.sct.subset, c("TBX18", "SHOX2", "BMP4", "ISL1", "TBX3", "NKX2-5"), pt.size = 0, cols = custom_colors$discrete, ncol = 2)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/SinusVenosus_Genes.Violin.UMAP.res.0.3.pdf", width = 10, height = 10)

VlnPlot(pacemaker.combined.sct.subset, c("TBX3", "ISL1", "BMP4", "CPNE5", "NKX2-5", "NPPA", "HAMP", "ADM"), group.by = "Annotation", pt.size = 0, cols = custom_colors$discrete, ncol = 2)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/SAN_Transitional_Genes.Violin.UMAP.res.0.3.pdf", width = 12, height = 13)
```

Find markers at Res 0.3.subset using RNA assay (which is after filtering)
```{r}
# find markers for every cluster compared to all remaining cells, report only the positive ones
DefaultAssay(pacemaker.combined.sct.subset) <- "RNA"
pacemaker.combined.sct.subset <- NormalizeData(pacemaker.combined.sct.subset, normalization.method = "LogNormalize", scale.factor = 10000)
pacemaker.combined.sct.subset[["RNA"]]@data

pacemaker.combined.sct.subset.markers <- FindAllMarkers(pacemaker.combined.sct.subset, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, assay = "RNA", slot = "data")

save(pacemaker.combined.sct.subset.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.3.subset_RNA_Assay.RData")

write.table(as.data.frame(pacemaker.combined.sct.subset.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.3.subset_RNA_Assay.txt", quote = F, sep = "\t")

pacemaker.combined.sct.subset.markers.fdr <- filter(pacemaker.combined.sct.subset.markers, p_val_adj < 0.01)

table(pacemaker.combined.sct.subset.markers.fdr$cluster)
#  0   1   2   3   4   5   6   7 
#178 163  69 594 284 702 941 572

pacemaker.combined.sct.subset.markers.fdr %>%
    group_by(cluster) %>%
    slice_head(n = 200) -> top200

write.table(as.data.frame(top200),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.3.subset_top200.txt", quote = F, sep = "\t")
####
#pacemaker.combined.sct.subset.markers.fdr$cluster = as.vector(as.numeric(pacemaker.combined.sct.subset.markers.fdr$cluster))
top200list <- matrix(0, nrow = 200, ncol = 8)
top200_df <- as.data.frame(top200)
for(i in seq(0,7,1)){
  j <- i+1
  top200list[1:length(top200_df[top200_df$cluster == i,7]),j] <- top200_df[top200_df$cluster == i,7]
}
colnames(top200list) <- paste0("C_", seq(0,7,1))

write.table(as.data.frame(top200list),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.3.subset_top200list.txt", quote = F, sep = "\t", row.names = F)


### Plot top 5 markers for each cluster based on avg_log2FC
pacemaker.combined.sct.subset.markers.fdr %>%
    group_by(cluster) %>%
    slice_max(avg_log2FC, n = 5) -> top5


#DotPlot(filtered.D34.samples.select, features = unique(tmp_top5_select$gene), cols= c("white","#9632B8"), cluster.idents = FALSE) + RotatedAxis() + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + xlab("Genes") + ylab("Clusters")
#ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/UMAP/Top5.markers_dotplot_select_clusters.pdf", width = 9, height = 5)

DotPlot_scCustom(pacemaker.combined.sct.subset, features = unique(top5$gene), flip_axes = T, remove_axis_titles = FALSE, colors_use = rev(viridis(1000))) + xlab("Genes") + ylab("Clusters")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Top5.markers_dotplot_subset_clusters_RNA_Assay_Res0.3.pdf", width = 6, height = 10)
```


Stacked Violin Plots
```{r}
VlnPlot(pacemaker.combined.sct.subset, features = c("TBX18", "SHOX2", "ISL1", "TBX3", "NKX2-5"), stack = TRUE, cols = paletteer_d("ggthemes::excel_Berlin")[1:5], flip = TRUE, idents = c("0", "1", "2", "3", "4", "5")) + theme(legend.position = "none") + ggtitle("Identity on y-axis")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Pacemaker_Genes.StackedViolin.Res0.3.pdf", width = 6, height = 5)

VlnPlot(pacemaker.combined.sct.subset, features = c("TNNT2", "ACTN2", "ALDH1A2", "WT1", "KRT8", "KRT18", "BNC1", "PDPN", "PHOX2B", "PRPH", "CHRNA3"), stack = TRUE, cols = paletteer_d("ggsci::planetexpress_futurama"), flip = TRUE, assay = "SCT", slot = "data") + theme(legend.position = "none") + ggtitle("Identity on y-axis")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Pan_cardiac_vs_non_cardio_Genes.StackedViolin.Res.0.3.pdf", width = 6, height = 6)

VlnPlot(pacemaker.combined.sct.subset, features = c("TBX3", "NPPA", "HAMP", "MYH6", "NKX2-5", "CPNE5", "ISL1", "BMP4"), stack = TRUE, cols = paletteer_d("ggthemes::few_Dark")[2:9], flip = TRUE, assay = "SCT", slot = "data") + theme(legend.position = "none") + ggtitle("Identity on y-axis")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/TransitionalCells_Genes.StackedViolinRes.0.3.pdf", width = 6, height = 6)

VlnPlot(pacemaker.combined.sct.subset, features = c("MYL7", "NPPA", "NR2F2", "MYL4"), stack = TRUE, cols = paletteer_d("ggthemes::wsj_rgby"), flip = TRUE, assay = "SCT", slot = "data") + theme(legend.position = "none") + ggtitle("Identity on y-axis")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/AtrialCells_Genes.StackedViolinRes.0.3.pdf", width = 6, height = 6)
```


Remove cluster 3 (high mitochondrial expression) and cluster 7 (stressed cardiomyocytes) at Res 0.2.
```{r}
pacemaker.combined.sct <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Clustered.pacemaker.combined.sct.rds")
#Idents(object = pacemaker.combined.sct) <- "integrated_snn_res.0.2"

pacemaker.combined.sct.subset <- subset(pacemaker.combined.sct, subset = integrated_snn_res.0.2 == "0" | integrated_snn_res.0.2 == "1" | integrated_snn_res.0.2 == "2" | integrated_snn_res.0.2 == "4" | integrated_snn_res.0.2 == "5" | integrated_snn_res.0.2 == "6" | integrated_snn_res.0.2 == "8"  | integrated_snn_res.0.2 == "9")

table(pacemaker.combined.sct.subset$integrated_snn_res.0.2)
pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset =  as.vector(pacemaker.combined.sct.subset$integrated_snn_res.0.2)
pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset=="4"]=3
pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset=="5"]=4
pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset=="6"]=5
pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset=="8"]=6
pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset[pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset=="9"]=7
table(pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset)

pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset = factor(pacemaker.combined.sct.subset$integrated_snn_res.0.2.subset, levels = 0:7)
Idents(object = pacemaker.combined.sct.subset) <- "integrated_snn_res.0.2.subset"
DimPlot(pacemaker.combined.sct.subset,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        #cols = custom_colors$discrete, 
        cols = paletteer_d("ggthemes::Tableau_10"), 
        pt.size = 1,
        label.size = 10)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/UMAP.res.0.2.subset.pdf", width = 10, height = 8)
```


Find markers at Res 0.2.subset using RNA assay.
```{r}
# find markers for every cluster compared to all remaining cells, report only the positive ones
DefaultAssay(pacemaker.combined.sct.subset) <- "RNA"
pacemaker.combined.sct.subset <- NormalizeData(pacemaker.combined.sct.subset, normalization.method = "LogNormalize", scale.factor = 10000)
pacemaker.combined.sct.subset[["RNA"]]@data

pacemaker.combined.sct.subset.markers <- FindAllMarkers(pacemaker.combined.sct.subset, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, assay = "RNA", slot = "data")

save(pacemaker.combined.sct.subset.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.2.subset_RNA_Assay.RData")

write.table(as.data.frame(pacemaker.combined.sct.subset.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.2.subset_RNA_Assay.txt", quote = F, sep = "\t")

pacemaker.combined.sct.subset.markers.fdr <- filter(pacemaker.combined.sct.subset.markers, p_val_adj < 0.01)

table(pacemaker.combined.sct.subset.markers.fdr$cluster)
# 0    1    2    3    4    5    6    7 
#  73  182  712  553  272  714 1002  576

pacemaker.combined.sct.subset.markers.fdr %>%
    group_by(cluster) %>%
    slice_head(n = 200) -> top200

write.table(as.data.frame(top200),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.2.subset_top200.txt", quote = F, sep = "\t")
####
#pacemaker.combined.sct.subset.markers.fdr$cluster = as.vector(as.numeric(pacemaker.combined.sct.subset.markers.fdr$cluster))
top200list <- matrix(0, nrow = 200, ncol = 8)
top200_df <- as.data.frame(top200)
for(i in seq(0,7,1)){
  j <- i+1
  top200list[1:length(top200_df[top200_df$cluster == i,7]),j] <- top200_df[top200_df$cluster == i,7]
}
colnames(top200list) <- paste0("C_", seq(0,7,1))

write.table(as.data.frame(top200list),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/FindAllMarkers_Res0.2.subset_top200list.txt", quote = F, sep = "\t", row.names = F)


### Plot top 5 markers for each cluster.
pacemaker.combined.sct.subset.markers.fdr %>%
    group_by(cluster) %>%
    slice_head(n = 5) -> top5


#DotPlot(filtered.D34.samples.select, features = unique(tmp_top5_select$gene), cols= c("white","#9632B8"), cluster.idents = FALSE) + RotatedAxis() + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + xlab("Genes") + ylab("Clusters")
#ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/UMAP/Top5.markers_dotplot_select_clusters.pdf", width = 9, height = 5)

DotPlot_scCustom(pacemaker.combined.sct.subset, features = unique(top5$gene), flip_axes = T, remove_axis_titles = FALSE, colors_use = rev(viridis(1000))) + xlab("Genes") + ylab("Clusters")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Top5.markers_dotplot_subset_clusters_RNA_Assay.pdf", width = 6, height = 10)
```


Stacked Violin Plots
```{r}
VlnPlot(pacemaker.combined.sct.subset, features = c("TBX18", "SHOX2", "ISL1", "TBX3", "NKX2-5"), stack = TRUE, cols = paletteer_d("ggthemes::excel_Berlin")[1:5], flip = TRUE, idents = c("0", "1", "2", "3", "4", "5")) + theme(legend.position = "none") + ggtitle("Identity on y-axis")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Pacemaker_Genes.StackedViolin.pdf", width = 6, height = 5)

VlnPlot(pacemaker.combined.sct.subset, features = c("TNNT2", "ACTN2", "ALDH1A2", "WT1", "KRT8", "KRT18", "BNC1", "PDPN", "PHOX2B", "PRPH", "CHRNA3"), stack = TRUE, cols = paletteer_d("ggsci::planetexpress_futurama"), flip = TRUE, assay = "SCT", slot = "data") + theme(legend.position = "none") + ggtitle("Identity on y-axis")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/Pan_cardiac_vs_non_cardio_Genes.StackedViolin.pdf", width = 6, height = 6)


VlnPlot(pacemaker.combined.sct.subset, features = c("TBX3", "NPPA", "HAMP", "MYH6", "NKX2-5", "CPNE5", "ISL1", "BMP4"), stack = TRUE, cols = paletteer_d("ggthemes::few_Dark")[2:9], flip = TRUE, assay = "SCT", slot = "data") + theme(legend.position = "none") + ggtitle("Identity on y-axis")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/TransitionalCells_Genes.StackedViolin.pdf", width = 6, height = 6)


VlnPlot(pacemaker.combined.sct.subset, features = c("MYL7", "NPPA", "NR2F2", "MYL4"), stack = TRUE, cols = paletteer_d("ggthemes::wsj_rgby"), flip = TRUE, assay = "SCT", slot = "data") + theme(legend.position = "none") + ggtitle("Identity on y-axis")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/UMAP/AtrialCells_Genes.StackedViolin.pdf", width = 6, height = 6)
```



Subclustering of SAN tail and head cells from Res 0.3
#########################################################################################################
#########################################################################################################
```{r}
pacemaker.combined.sct <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/Clustered.pacemaker.combined.sct.rds")
load("/Data/iPSC_pacemaker/Seurat/D34_First/D34.filtered.samples.RData")

tail.head.cells <- filter(pacemaker.combined.sct@meta.data, integrated_snn_res.0.3 == "1" | integrated_snn_res.0.3 == "2")

# Extract expression matrix for fibroblasts.
tail.head.raw.data <- as.matrix(GetAssayData(filtered.samples, slot = "counts")[, rownames(tail.head.cells)])
# Create a Seurat object for fibroblast.
tail.head <- CreateSeuratObject(tail.head.raw.data, project = "tail.head", meta.data = tail.head.cells)
#table(colnames(tail.head)==rownames(tail.head.cells))
```

Evaluating effects of cell cycle.
```{r}
# Normalize the counts
seurat_phase <- NormalizeData(tail.head)
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
s.genes[s.genes=="MLF1IP"] = "CENPU"
g2m.genes[g2m.genes=="FAM64A"] = "PIMREG"
g2m.genes[g2m.genes=="HN1"] = "JPT1"
seurat_phase <- CellCycleScoring(seurat_phase, 
                                 g2m.features = g2m.genes, 
                                 s.features = s.genes)
# Calculate the percentage of cells in different phases.
phase_count <- table(seurat_phase@meta.data$orig.ident, seurat_phase@meta.data$Phase)
phase_per <- prop.table(table(seurat_phase@meta.data$orig.ident, seurat_phase@meta.data$Phase), 1)*100
write.table(phase_count, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Covariates/CellCyclePhaseCount.txt", quote = F, sep = "\t")
write.table(phase_per, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Covariates/CellCyclePhasePercentage.txt", quote = F, sep = "\t")
# Identify the most variable genes
seurat_phase <- FindVariableFeatures(seurat_phase, 
                     selection.method = "vst",
                     nfeatures = 4000, 
                     verbose = FALSE)
# Scale the counts
seurat_phase <- ScaleData(seurat_phase)
# Perform PCA
seurat_phase <- RunPCA(seurat_phase, ndims.print = 1:40, nfeatures.print = 10)
# Run UMAP
seurat_phase <- RunUMAP(seurat_phase, dims = 1:40)
# Plot the UMAP colored by cell cycle phase
DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "Phase",
        split.by = "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Covariates/SplitCellCyclePlot.pdf")
DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Covariates/MergedCellCyclePlot.pdf")

DimPlot(seurat_phase,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        group.by= "orig.ident") + scale_color_discrete(limits = c("D34_r0_P", "D34_r1_P", "D34_r2_P"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Covariates/MergedCellCyclePlot_Samples.pdf")
```
We do see large differences due to cell cycle. Based on this plot, we would regress out the variation due to cell cycle.

Add cell cycle score to tail.head for SCTransform.
```{r}
table(rownames(seurat_phase@meta.data)==rownames(tail.head@meta.data))
tail.head@meta.data <- mutate(tail.head@meta.data, S.Score = seurat_phase@meta.data$S.Score, G2M.Score = seurat_phase@meta.data$G2M.Score, Phase = seurat_phase@meta.data$Phase)
## Keep only the useful info.
tail.head@meta.data <- select(tail.head@meta.data, 1:8)
```

Apply sctransform normalization while regress out cell cycle scoring.
```{r}
tail.head <- SCTransform(tail.head, method = "glmGamPoi", vars.to.regress = c("G2M.Score", "S.Score"), variable.features.n = 3000)
# Save the seurat object
saveRDS(tail.head, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/SCTransform.tail.head.cells.rds")
```

```{r}
tail.head <- RunPCA(tail.head)
DimPlot(tail.head, reduction = "pca", group.by = "orig.ident", cols = custom_colors$discrete) + scale_color_discrete(limits = c("D34_r0_P", "D34_r1_P", "D34_r2_P"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/PCs/MergedPCAPlot.pdf")

tail.head$orig.ident <- factor(tail.head$orig.ident, levels = unique(tail.head$orig.ident))

DimPlot(tail.head, reduction = "pca", split.by = "orig.ident", group.by = "orig.ident", cols = custom_colors$discrete) + scale_color_discrete(limits = c("D34_r0_P", "D34_r1_P", "D34_r2_P"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/PCs/SplitPCAPlot.pdf", width = 30, height = 5)

# Plot the elbow plot
ElbowPlot(object = tail.head, ndims = 50)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/PCs/ElbowPlot.pdf")
```

Cluster the cells
```{r}
# Determine the K-nearest neighbor graph
tail.head <- FindNeighbors(object = tail.head, 
                                dims = 1:40)
                                
# Determine the clusters for various resolutions                                
tail.head <- FindClusters(object = tail.head,
                               resolution = c(0.2))
# Look at cluster IDs of the first 5 cells
head(Idents(tail.head), 5)
table(as.vector(tail.head$SCT_snn_res.0.2))
```

Run non-linear dimensional reduction (UMAP/tSNE)
```{r}
tail.head <- RunUMAP(tail.head, dims = 1:40)
# Plot the UMAP
# Assign identity of clusters
Idents(object = tail.head) <- "SCT_snn_res.0.2"
DimPlot(tail.head,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/UMAP/UMAP.res.0.2.pdf", width = 10, height = 8)

# Extract identity and sample information from seurat object to determine the number of cells per cluster per sample
n_cells <- FetchData(tail.head, 
                     vars = c("ident", "orig.ident")) %>%
        dplyr::count(ident, orig.ident) %>%
        tidyr::spread(ident, n)

write.table(n_cells,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/UMAP/N_Cells_For_Each_Clusters.txt", quote = F, sep = "\t", row.names = F)

# UMAP of cells in each cluster by sample
DimPlot(tail.head, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/UMAP/SplitSamples.UMAP.res.0.2.pdf", width = 15, height = 5)

DimPlot(tail.head, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/UMAP/MergedSamples.UMAP.res.0.2.diff.colors.pdf", width = 10, height = 8)
```
Batch effect exists, so decide to do integration.

```{r}
library(harmony)
tail.head <- RunHarmony(tail.head, "orig.ident")
tail.head <- RunUMAP(tail.head, reduction = "harmony", dims = 1:40)
tail.head <- FindNeighbors(tail.head, reduction = "harmony", dims = 1:40)
tail.head <- FindClusters(object = tail.head, resolution = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

Idents(object = tail.head) <- "SCT_snn_res.0.2"
DimPlot(tail.head,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/UMAP.res.0.2.tail.head.cells.3000.features.pdf", width = 10, height = 8)

DimPlot(tail.head,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        group.by = "SCT_snn_res.0.3",
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/UMAP.res.0.3.tail.head.cells.3000.features.pdf", width = 10, height = 8)

#Segregation of clusters by sample
# UMAP of cells in each cluster by sample
DimPlot(tail.head, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/SplitSamples.UMAP.res.0.2.tail.head.cells.top.3000.features.pdf", width = 15, height = 5)

DimPlot(tail.head, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/MergedSamples.UMAP.res.0.2.tail.head.cells.top.3000.features.pdf", width = 10, height = 8)

DimPlot(tail.head, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/MergedSamples.UMAP.res.0.2.diff.colors.tail.head.cells.top.3000.features.pdf", width = 10, height = 8)


DimPlot(tail.head, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        pt.size = 1,
        label.size = 8,
        group.by = "SCT_snn_res.0.3")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/UMAP.res.0.3.diff.colors.tail.head.cells.top.3000.features.pdf", width = 10, height = 8)


# Count the number of clusters at different resolutions.
tail.head@meta.data = select(tail.head@meta.data, c(1:10,13:22,11,23:34))
Ncluster <- c()
for(res in c(11:33)){
  Ncluster <- c(Ncluster, length(unique(as.vector(tail.head@meta.data[,res]))))
#unique(pacemarker.cells@meta.data$SCT_snn_res.2)
}
res_ncluster <- data.frame(Res = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0), Ncluster = Ncluster)
write.table(res_ncluster,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/ResolutionVsNclusters.txt", quote = F, sep = "\t", row.names = F)
```

Export rds file for BBrowser use.
```{r}
#require(rBCS)
#ExportSeurat(tail.head, "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3.head/Harmony.tail.head.bcs")
saveRDS(tail.head, file = "/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony.tail.head.rds")
```

Silhouette plot
```{r}
distance_matrix <- dist(Embeddings(tail.head[['pca']])[, 1:40])

sil_score_matrix <- matrix(NA, nrow=dim(tail.head@meta.data)[1], ncol=16)
tail.head@meta.data <- select(tail.head@meta.data, 1:10, 18:33)
for(i in 11:26){
  clusters <- tail.head@meta.data[,i]
  silhouette <- silhouette(as.numeric(clusters), dist = distance_matrix)
  sil_score_matrix[,i-10] <- silhouette[,3]
}

colnames(sil_score_matrix) = paste0("silhouette_score_res.", c(0.08, 0.09, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

tail.head@meta.data <- cbind(tail.head@meta.data, sil_score_matrix)

tmp_score <- select(tail.head@meta.data, 27:42)
library(reshape2)
long_tmp_score <- melt(tmp_score)

long_tmp_score$variable = str_replace(long_tmp_score$variable, "silhouette_score_res.", "")

library(ggpubr)
ggviolin(long_tmp_score, "variable", "value", fill = "variable",
   palette = custom_colors$discrete,
   add = "boxplot", add.params = list(fill = "white"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/silhouette_violin_plot.pdf", width = 10, height = 5)
```

Exploring known cell type markers
```{r}
# Pacemaker cell genes
FeaturePlot(tail.head, 
            reduction = "umap", 
            features = c("TBX18", "SHOX2", "ISL1", "NKX2-5", "TBX3"), 
            order = FALSE,
            #min.cutoff = 'q10', 
            label = TRUE)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/tail.headCell_Genes.UMAP.pdf", width = 9, height = 10)

# Violin plot
VlnPlot(tail.head, c("TBX18", "SHOX2", "ISL1", "NKX2-5", "TBX3"), group.by = "orig.ident", pt.size = 0, cols = custom_colors$discrete, ncol = 4)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/tail.headCell_Genes.Violin.UMAP.pdf", width = 16, height = 5)
```

```{r}
DotPlot(tail.head, features = rev(c("NKX2-5", "TNNT2", "MYH6", "MYH7","TBX18", "SHOX2", "ISL1", "TBX3", "HCN4", "KCNJ3", "GJD3", "MYL2", "HEY2", "IRX4", "MYL7", "NPPA", "NR2F2", "MSX2", "TBX2")), cols= "Spectral", group.by = "orig.ident") + RotatedAxis() + coord_flip()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/James_Markers_DotPlot.UMAP.pdf", width = 8, height = 6)
```

```{r}
# Single cell heatmap of feature expression
DoHeatmap(tail.head, features = c("NKX2-5", "TNNT2", "MYH6", "MYH7","TBX18", "SHOX2", "ISL1", "TBX3", "HCN4", "KCNJ3", "GJD3", "MYL2", "HEY2", "IRX4", "MYL7", "NPPA", "NR2F2", "MSX2", "TBX2"), size = 3, group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/Harmony/James_Markers_Heatmap.UMAP.res.0.2.pdf")
```

Cross check the clusters with clustering of D34 packmaker cell culture.
```{r}
tmp1 <- select(tail.head@meta.data,13:16)
tmp1 <- mutate(tmp1, Cells =  rownames(tmp1))
tmp2 <- select(pacemaker.combined.sct@meta.data,21:33)
tmp2 <- mutate(tmp2, Cells =  rownames(tmp2))
tmp <- left_join(tmp1, tmp2)
table(tmp$SCT_snn_res.0.3, tmp$integrated_snn_res.0.3)
  
#       0    1    2    3    4    5    6    7    8    9   10   11   12   13
#  0    0 3178   36    0    0    0    0    0    0    0    0    0    0    0
#  1    0  237 2317    0    0    0    0    0    0    0    0    0    0    0
#  2    0  136 1878    0    0    0    0    0    0    0    0    0    0    0
#  3    0  747   46    0    0    0    0    0    0    0    0    0    0    0
```

Cross check the clusters with clustering of timecourse dataset.
```{r}
filtered.pacemaker.samples <- readRDS("/Data/iPSC_pacemaker/Seurat/Pacemaker_Cells/Second_Round/Clustered.filtered.pacemaker.samples.rds")
Idents(object = filtered.pacemaker.samples) <- "SCT_snn_res.0.6"
tmp1 <- select(filtered.pacemaker.samples@meta.data,14,16)
tmp1 <- mutate(tmp1, Cells =  rownames(tmp1))
tmp2 <- select(pacemaker.combined.sct@meta.data,21:33)
tmp2 <- mutate(tmp2, Cells =  rownames(tmp2))
tmp <- left_join(tmp2, tmp1)
table(tmp$SCT_snn_res.0.6, tmp$integrated_snn_res.0.3)
```

Differential expression analysis using assay "RNA" slot "data" as recommended by satijalab. "https://github.com/satijalab/seurat/discussions/4000".  https://github.com/satijalab/seurat/discussions/4032.
```{r}
## Try to use RNA assay for differential expression analysis.
tail.head.RNA <- tail.head
Idents(object = tail.head.RNA) <- "SCT_snn_res.0.3"
DefaultAssay(tail.head.RNA) <- "RNA"
tail.head.RNA <- NormalizeData(tail.head.RNA, normalization.method = "LogNormalize", scale.factor = 10000)
tail.head.RNA[["RNA"]]@data

tail_vs_head_RNA <- FindMarkers(tail.head.RNA, ident.1 = c("1","2","3"), ident.2 = "0", min.pct = 0.25, assay = "RNA", slot = "data")
write.table(tail_vs_head_RNA,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/DE_Tail_vs_Head_RNA_Assay.txt", quote = F, sep = "\t", row.names = T)
dim(tail_vs_head_RNA)
#[1] 413   5
tail_vs_head_RNA_Filter <- filter(tail_vs_head_RNA, abs(avg_log2FC) > 0.25 & p_val_adj < 0.05)
dim(tail_vs_head_RNA_Filter)
#[1] 412   5
dim(filter(tail_vs_head_RNA_Filter, avg_log2FC > 0))
#[1] 133   5
######################################################################################################################################################
##### Try DESeq2 for differential expression. Takes very long to run.
tail_vs_head_RNA_DESeq2 <- FindMarkers(tail.head.RNA, ident.1 = c("1","2","3"), ident.2 = "0", min.pct = 0.25, assay = "RNA", slot = "data", test.use = "DESeq2")
write.table(tail_vs_head_RNA_DESeq2,"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/DE_Tail_vs_Head_RNA_Assay_DESeq2.txt", quote = F, sep = "\t", row.names = T)
###### DESeq2 report more genes than others based on logfc 0.25 and padj 0.05. 95% of the wilcox genes are in DESeq2's DEGs..
#######################################################################################################################################################
```


Plot genes from differential expression testing between clusters.
```{r}
tail_vs_head_RNA_Filter <- mutate(tail_vs_head_RNA_Filter, Gene = rownames(tail_vs_head_RNA_Filter))

# Single cell heatmap of feature expression
tail.head.RNA.subset <- subset(tail.head.RNA, subset = SCT_snn_res.0.3 == "0" | SCT_snn_res.0.3 == "1" | SCT_snn_res.0.3 == "2" | SCT_snn_res.0.3 == "3")
mat<- tail.head.RNA.subset[["RNA"]]@data[rownames(tail_vs_head_RNA_Filter),] %>% as.matrix()
#tail.head.RNA.subset[["SCT"]]@data[unique(rownames(combine_DEGs)),]
## scale the rows
mat<- t(scale(t(mat)))
cluster_anno<- tail.head.RNA.subset@meta.data$SCT_snn_res.0.3
### Annotate certain genes.
ha = rowAnnotation(foo = anno_mark(at = c(2,9,29,35,53,62,65,100,89,109,138,141,165,214,225), labels = rownames(tail_vs_head_RNA_Filter)[c(2,9,29,35,53,62,65,100,89,109,138,141,165,214,225)]))

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/tail.head_DEGs_heatmap.pdf", height = 8, width = 10)
ht <- Heatmap(mat, name = "Expression",  
        column_split = factor(cluster_anno),
        cluster_columns = FALSE,
        show_column_dend = FALSE,
        cluster_column_slices = TRUE,
        column_title_gp = gpar(fontsize = 8),
        column_gap = unit(0, "mm"),
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        #col = col_fun,
        row_names_gp = gpar(fontsize = 4),
        column_title_rot = 0,
        #top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = scales::hue_pal()(7)))),
        #HeatmapAnnotation(bar = cluster_anno, col = list(bar = c("ACM" = "red", "VCM" = "green"))),
        top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = paletteer_d("ggsci::planetexpress_futurama")[c(1,2,3,4)]))),
        #right_annotation = HeatmapAnnotation(foo = anno_barplot(ACM_vs_VCM_DE_order$avg_log2FC, baseline = 0)),
        show_column_names = FALSE,
        show_row_names = FALSE,
        use_raster = TRUE,
        raster_quality = 4,
        right_annotation = ha,
        col = colorRamp2(c(-4, 0, 4), c("#56B4E9", "#F7FAFF", "#E69F00")))
draw(ht)
dev.off()
```


```{r}
# Plot top 25 pathways
pathways <- read.table("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/DEGs_Metascape.txt", sep = "\t", quote = "", header = T, comment.char = "", check.names = F)

## Plot FDR values rather than p-values for GO terms.
pathways_fdr <- as.data.frame(apply(pathways[1:6], 2, function(x){log10(p.adjust(10^x))}))
pathways_fdr <- mutate(pathways_fdr, Category = pathways$Category, Description = pathways$Description)
GO_fdr <- filter(pathways_fdr, Category == "GO Biological Processes")

### C0_vs_C1
C0_vs_C1_GO_Up_fdr <- GO_fdr %>% arrange(C0_vs_C1_Up) %>% head(n=10)
C0_vs_C1_GO_Down_fdr <- GO_fdr %>% arrange(C0_vs_C1_Down) %>% head(n=10)
C0_vs_C1_Up_Down_top10_fdr <- GO_fdr %>% filter(Description %in% unique(c(C0_vs_C1_GO_Up_fdr$Description, C0_vs_C1_GO_Down_fdr$Description))) %>% select(C0_vs_C1_Down, C0_vs_C1_Up, Description) %>% arrange(desc(C0_vs_C1_Up))
C0_vs_C1_Up_Down_top10_long_fdr <- melt(C0_vs_C1_Up_Down_top10_fdr)
C0_vs_C1_Up_Down_top10_long_fdr$variable <- factor(C0_vs_C1_Up_Down_top10_long_fdr$variable, levels = c("C0_vs_C1_Up","C0_vs_C1_Down"))
C0_vs_C1_Up_Down_top10_long_fdr$Description <- factor(C0_vs_C1_Up_Down_top10_long_fdr$Description, levels = unique(C0_vs_C1_Up_Down_top10_fdr$Description))
ggplot(C0_vs_C1_Up_Down_top10_long_fdr,aes(x=variable, y=Description)) + geom_point(aes(size=-value), colour = "#6BA3D6") + xlab("DEG Sets") + ylab("GO Pathways") + labs(size="-LogFDR") + theme(axis.text=  element_text(color = "black"), axis.title =  element_text(size = 9))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C1_Up_Down_Top_10_Pathways_fdr.pdf", width = 6, height = 8)

# Only plot top 10 pathways in up DEGs.
C0_vs_C1_GO_Up_fdr$Description = factor(C0_vs_C1_GO_Up_fdr$Description, levels = rev(C0_vs_C1_GO_Up_fdr$Description))
ggplot(C0_vs_C1_GO_Up_fdr, aes(x = -C0_vs_C1_Up, y=Description)) + geom_bar(stat="identity", fill="#A39FC9") + xlab("-log10(FDR)") + ylab("GO Biological Processes") + theme_classic()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C1_Up_Top_10_Pathways_fdr.pdf", width = 6, height = 6)
# Only plot top 10 pathways in Down DEGs.
C0_vs_C1_GO_Down_fdr$Description = factor(C0_vs_C1_GO_Down_fdr$Description, levels = rev(C0_vs_C1_GO_Down_fdr$Description))
ggplot(C0_vs_C1_GO_Down_fdr, aes(x = -C0_vs_C1_Down, y=Description)) + geom_bar(stat="identity", fill="#92AA4C") + xlab("-log10(FDR)") + ylab("GO Biological Processes") + theme_classic()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Integration/DE/C0_vs_C1_Down_Top_10_Pathways_fdr.pdf", width = 6, height = 6)
```

Find markers at Res 0.3 using RNA assay.
```{r}
tail.head.RNA <- tail.head
Idents(object = tail.head.RNA) <- "SCT_snn_res.0.3"
DefaultAssay(tail.head.RNA) <- "RNA"
tail.head.RNA <- NormalizeData(tail.head.RNA, normalization.method = "LogNormalize", scale.factor = 10000)
tail.head.RNA[["RNA"]]@data

tail.head.RNA.markers <- FindAllMarkers(tail.head.RNA, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, assay = "RNA", slot = "data")

save(tail.head.RNA.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/FindAllMarkers_Res0.3_RNA_Assay.RData")

write.table(as.data.frame(tail.head.RNA.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/FindAllMarkers_Res0.3_RNA_Assay.txt", quote = F, sep = "\t")

tail.head.RNA.markers.fdr <- filter(tail.head.RNA.markers, p_val_adj < 0.01)

table(tail.head.RNA.markers.fdr$cluster)
#  0   1   2   3 
#279  64  71 115 

tail.head.RNA.markers.fdr %>%
    group_by(cluster) %>%
    slice_head(n = 200) -> top200

write.table(as.data.frame(top200),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/FindAllMarkers_Res0.3.RNA_top200.txt", quote = F, sep = "\t")

top200list <- matrix(0, nrow = 200, ncol = 4)
top200_df <- as.data.frame(top200)
for(i in seq(0,3,1)){
  j <- i+1
  top200list[1:length(top200_df[top200_df$cluster == i,7]),j] <- top200_df[top200_df$cluster == i,7]
}
colnames(top200list) <- paste0("C_", seq(0,3,1))

write.table(as.data.frame(top200list),"/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/FindAllMarkers_Res0.3.RNA_top200list.txt", quote = F, sep = "\t", row.names = F)


### Plot top 5 markers for each cluster.
tail.head.RNA.markers.fdr %>%
    group_by(cluster) %>%
    slice_head(n = 5) -> top5

DotPlot_scCustom(tail.head.RNA, features = unique(top5$gene), flip_axes = T, remove_axis_titles = FALSE, colors_use = rev(viridis(1000))) + xlab("Genes") + ylab("Clusters")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/Top5.markers_dotplot_Res.0.3.RNA_clusters_RNA_Assay.pdf", width = 6, height = 10)

### Plot top 10 markers for each cluster.
tail.head.RNA.markers.fdr %>%
    group_by(cluster) %>%
    slice_head(n = 10) -> top10

DotPlot_scCustom(tail.head.RNA, features = unique(top10$gene), flip_axes = T, remove_axis_titles = FALSE, colors_use = rev(viridis(1000))) + xlab("Genes") + ylab("Clusters")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/PacemakerCells/Tail_Head_Res0.3/DE/Top10.markers_dotplot_Res.0.3.RNA_clusters_RNA_Assay.pdf", width = 6, height = 10)
```
