---
title: "D34.Pacemaker.Cells"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(RColorBrewer)
library(scDblFinder)
library(SummarizedExperiment)
library(stringr)
library(tidyr)
library(ggforce)
library(cluster)
library(rBCS)
library(reshape2)
library(ggpubr)
```


```{r}
# Define colors
custom_colors <- list()
colors_dutch <- c(
  '#FFC312','#C4E538','#12CBC4','#FDA7DF','#ED4C67',
  '#F79F1F','#A3CB38','#1289A7','#D980FA','#B53471',
  '#EE5A24','#009432','#0652DD','#9980FA','#833471',
  '#EA2027','#006266','#1B1464','#5758BB','#6F1E51'
)

colors_spanish <- c(
  '#40407a','#706fd3','#f7f1e3','#34ace0','#33d9b2',
  '#2c2c54','#474787','#aaa69d','#227093','#218c74',
  '#ff5252','#ff793f','#d1ccc0','#ffb142','#ffda79',
  '#b33939','#cd6133','#84817a','#cc8e35','#ccae62'
)

custom_colors$discrete <- c(colors_dutch, colors_spanish)
```


```{r}
load("/Data/iPSC_pacemaker/Seurat/D34_First/D34.filtered.samples.RData")
ventricular.cells <- subset(filtered.samples, subset = orig.ident == "D34_r0_V" | orig.ident == "D34_r1_V" | orig.ident == "D34_r2_V")
ventricular.cells
#An object of class Seurat 
#27519 features across 20712 samples within 1 assay 
#Active assay: RNA (27519 features, 0 variable features)
```

```{r}
# Normalize the counts
seurat_phase <- NormalizeData(ventricular.cells)
```

Evaluating effects of cell cycle.
```{r}
# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

# Manually change the names of these cell cycle genes.
s.genes[s.genes=="MLF1IP"] = "CENPU"
g2m.genes[g2m.genes=="FAM64A"] = "PIMREG"
g2m.genes[g2m.genes=="HN1"] = "JPT1"

# Redo score cells for cell cycle
seurat_phase <- CellCycleScoring(seurat_phase, 
                                 g2m.features = g2m.genes, 
                                 s.features = s.genes)
# Calculate the percentage of cells in different phases.
phase_count <- table(seurat_phase@meta.data$orig.ident, seurat_phase@meta.data$Phase)
phase_per <- prop.table(table(seurat_phase@meta.data$orig.ident, seurat_phase@meta.data$Phase), 1)*100
write.table(phase_count, "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/CellCyclePhaseCount.txt", quote = F, sep = "\t")
write.table(phase_per, "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/CellCyclePhasePercentage.txt", quote = F, sep = "\t")

# Visualize the distribution of cell cycle markers across
RidgePlot(seurat_phase, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), group.by = "Phase", ncol = 2) & scale_fill_manual(values = custom_colors$discrete)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/Cell.cycle.markers.by.phase.pdf")

RidgePlot(seurat_phase, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), group.by = "orig.ident", ncol = 2) & scale_y_discrete(limits = rev(c("D34_r0_V", "D34_r1_V", "D34_r2_V"))) & scale_fill_manual(values = custom_colors$discrete)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/Cell.cycle.markers.by.sample.pdf")
```

```{r}
# Identify the most variable genes
seurat_phase <- FindVariableFeatures(seurat_phase, 
                     selection.method = "vst",
                     nfeatures = 4000, 
                     verbose = FALSE)
		     
# Scale the counts
seurat_phase <- ScaleData(seurat_phase)

# Perform PCA
seurat_phase <- RunPCA(seurat_phase, ndims.print = 1:40, nfeatures.print = 10)

# Run UMAP
seurat_phase <- RunUMAP(seurat_phase, dims = 1:40)

# Plot the UMAP colored by cell cycle phase
DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "Phase",
        split.by = "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/SplitCellCyclePlot.pdf", width = 10, height = 5)

DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/MergedCellCyclePlot.pdf")

DimPlot(seurat_phase,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        group.by= "orig.ident") + scale_color_discrete(limits = c("D34_r0_V", "D34_r1_V", "D34_r2_V"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/MergedCellCyclePlot_Samples.pdf")

#DimHeatmap(seurat_phase, dims = c(8, 10))
```
We do see large differences due to cell cycle. Based on this plot, we would regress out the variation due to cell cycle.

Evaluating effects of mitochodrial expression.
```{r}
# Check quartile values
summary(seurat_phase@meta.data$percent.mt)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#  0.000   8.002  11.579  13.278  16.507  49.961 
# Turn mitoRatio into categorical factor vector based on quartile values
seurat_phase@meta.data$mitoFr <- cut(seurat_phase@meta.data$percent.mt, 
                   breaks=c(-Inf, 8.002, 11.579, 16.507, Inf), 
                   labels=c("Low","Medium","Medium high", "High"))

# Plot the PCA colored by mitoFr
DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "mitoFr",
        split.by = "mitoFr")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/SplitMitoFrPlot.pdf")

DimPlot(seurat_phase,
        reduction = "umap",
        group.by= "mitoFr")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Covariates/MergedMitoFrPlot.pdf")
```
We do not see large differences due to mitochondrial percentage for the same Tissue. Based on this plot, we would not regress out the variation due to mitochondrial percentage.

Add cell cycle score to ventricular.cells for SCTransform.
```{r}
table(rownames(seurat_phase@meta.data)==rownames(ventricular.cells@meta.data))
ventricular.cells@meta.data <- mutate(ventricular.cells@meta.data, S.Score = seurat_phase@meta.data$S.Score, G2M.Score = seurat_phase@meta.data$G2M.Score, Phase = seurat_phase@meta.data$Phase)
```

Apply sctransform normalization while regress out cell cycle scoring.
```{r}
# SCTranform
## Adjust the limit for allowable object sizes within R (Default is 500 * 1024 ^ 2 = 500 Mb) using the following code:
#options(future.globals.maxSize = 10000 * 1024^2)
ventricular.cells <- SCTransform(ventricular.cells, method = "glmGamPoi", vars.to.regress = c("G2M.Score", "S.Score"), variable.features.n = 5000)

# Save the seurat object
#saveRDS(ventricular.cells, "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/SCTransform.ventricular.cells.rds")

#filtered.samples <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/SCTransform.filtered.samples.rds")
```


```{r}
ventricular.cells <- RunPCA(ventricular.cells)
DimPlot(ventricular.cells, reduction = "pca", group.by = "orig.ident", cols = custom_colors$discrete) + scale_color_discrete(limits = c("D34_r0_V", "D34_r1_V", "D34_r2_V"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/PCs/MergedPCAPlot.pdf")

ventricular.cells$orig.ident <- factor(ventricular.cells$orig.ident, levels = unique(ventricular.cells$orig.ident))

DimPlot(ventricular.cells, reduction = "pca", split.by = "orig.ident", group.by = "orig.ident", cols = custom_colors$discrete) + scale_color_discrete(limits = c("D34_r0_V", "D34_r1_V", "D34_r2_V"))
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/PCs/SplitPCAPlot.pdf", width = 30, height = 5)
```

Determining how many PCs to include in the clustering step to ensure that we are capturing the majority of the variation, or cell types, present in our dataset.
```{r}
# Explore heatmap of PCs
pdf("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/PCs/HeatmapOfPCs_1_12.pdf", width = 25, height = 15)
#DimHeatmap(ventricular.cells, dims = 1:25, cells = 500, balanced = TRUE, fast = FALSE)
DimHeatmap(ventricular.cells, dims = 1:12, cells = 500, balanced = TRUE)
dev.off()

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/PCs/HeatmapOfPCs_13_24.pdf", width = 25, height = 15)
DimHeatmap(ventricular.cells, dims = 13:24, cells = 500, balanced = TRUE)
dev.off()

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/PCs/HeatmapOfPCs_25_36.pdf", width = 25, height = 15)
DimHeatmap(ventricular.cells, dims = 25:36, cells = 500, balanced = TRUE)
dev.off()

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/PCs/HeatmapOfPCs_37_48.pdf", width = 25, height = 15)
DimHeatmap(ventricular.cells, dims = 37:48, cells = 500, balanced = TRUE)
dev.off()

pdf("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/PCs/HeatmapOfPCs_49_50.pdf", width = 25, height = 15)
DimHeatmap(ventricular.cells, dims = 49:50, cells = 500, balanced = TRUE)
dev.off()

# Plot the elbow plot
ElbowPlot(object = ventricular.cells, ndims = 50)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/PCs/ElbowPlot.pdf")
```
Decision: Use the first 40 PCs to generate the clusters.

Cluster the cells
```{r}
# Determine the K-nearest neighbor graph
ventricular.cells <- FindNeighbors(object = ventricular.cells, 
                                dims = 1:40)
                                
# Determine the clusters for various resolutions                                
ventricular.cells <- FindClusters(object = ventricular.cells,
                               resolution = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

# Look at cluster IDs of the first 5 cells
head(Idents(ventricular.cells), 5)

#ventricular.cells@meta.data <- select(ventricular.cells@meta.data, c(1,25,2:10,26:34,11:24))


# Count the number of clusters at different resolutions.
Ncluster <- c()
for(res in c(12:33)){
  Ncluster <- c(Ncluster, length(unique(as.vector(ventricular.cells@meta.data[,res]))))
#unique(ventricular.cells@meta.data$SCT_snn_res.2)
}
res_ncluster <- data.frame(Res = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0), Ncluster = Ncluster)
write.table(res_ncluster,"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/ResolutionVsNclusters.txt", quote = F, sep = "\t", row.names = F)
```

Run non-linear dimensional reduction (UMAP/tSNE)
```{r}
ventricular.cells <- RunUMAP(ventricular.cells, dims = 1:40)

# Plot the UMAP
# Assign identity of clusters
Idents(object = ventricular.cells) <- "SCT_snn_res.0.6"
DimPlot(ventricular.cells,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/UMAP/UMAP.res.0.6.pdf", width = 10, height = 8)
```


Segregation of clusters by sample
```{r}
# Extract identity and sample information from seurat object to determine the number of cells per cluster per sample
n_cells <- FetchData(ventricular.cells, 
                     vars = c("ident", "orig.ident")) %>%
        dplyr::count(ident, orig.ident) %>%
        tidyr::spread(ident, n)

write.table(n_cells,"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/UMAP/N_Cells_For_Each_Clusters.txt", quote = F, sep = "\t", row.names = F)

# UMAP of cells in each cluster by sample
DimPlot(ventricular.cells, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/UMAP/SplitSamples.UMAP.res.0.6.pdf", width = 15, height = 5)

DimPlot(ventricular.cells, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/UMAP/MergedSamples.UMAP.res.0.6.pdf", width = 10, height = 8)

DimPlot(ventricular.cells, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.pdf", width = 10, height = 8)

```
Using the top 5000 variable genes among D34 pacemaker cells shows batch effect. This is inconsistent with the analysis using all the D34 cells including pacemaker, atrial and ventricular cells, which also used top 5000 variable genes. However, these two set of 5000 variable genes are different. The reason is that the genes leading to batch effect in D34 pacemaker cell clustering is not included in the D34 all cell type clustering as difference among cell types dominate the difference between batches in D34 pacemaker cells.


Test to use top 3000 genes
```{r}
ventricular.cells.3000 <- SCTransform(ventricular.cells, method = "glmGamPoi", vars.to.regress = c("G2M.Score", "S.Score"), variable.features.n = 3000)

ventricular.cells.3000 <- RunPCA(ventricular.cells.3000)

# Determine the K-nearest neighbor graph
ventricular.cells.3000 <- FindNeighbors(object = ventricular.cells.3000, 
                                dims = 1:40)
                                
# Determine the clusters for various resolutions                                
ventricular.cells.3000 <- FindClusters(object = ventricular.cells.3000,
                               resolution = c(0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

ventricular.cells.3000 <- RunUMAP(ventricular.cells.3000, dims = 1:40)

# Plot the UMAP
# Assign identity of clusters
Idents(object = ventricular.cells.3000) <- "SCT_snn_res.0.6"

# UMAP of cells in each cluster by sample
DimPlot(ventricular.cells.3000, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/UMAP/SplitSamples.UMAP.res.0.6.3000.pdf", width = 15, height = 5)

DimPlot(ventricular.cells.3000, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/UMAP/MergedSamples.UMAP.res.0.6.3000.pdf", width = 10, height = 8)

DimPlot(ventricular.cells.3000, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.3000.pdf", width = 10, height = 8)

#length(intersect(rownames(filtered.D34.samples[["SCT"]]@scale.data), rownames(ventricular.cells.3000[["SCT"]]@scale.data)))
#[1] 2762
```
Less batch effect was observed using top 3000 genes.


#####################################################################################################################
#####################################################################################################################
#####################################################################################################################
Decide to do integration for the 3 replicates to see if batch effect is still observed using top 5000 variable genes.
```{r}
# split the dataset into a list of three seurat objects (each for one sample)
ventricular.list <- SplitObject(ventricular.cells, split.by = "orig.ident")

# normalize and identify variable features for each dataset independently
ventricular.list <- lapply(X = ventricular.list, FUN = function(x) {
    x <- SCTransform(x, method = "glmGamPoi", vars.to.regress = c("G2M.Score", "S.Score"), variable.features.n = 5000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = ventricular.list, nfeatures = 5000)
ventricular.list <- PrepSCTIntegration(object.list = ventricular.list, anchor.features = features)

# Perform integration
ventricular.anchors <- FindIntegrationAnchors(object.list = ventricular.list, normalization.method = "SCT", anchor.features = features)
ventricular.combined.sct <- IntegrateData(anchorset = ventricular.anchors, normalization.method = "SCT")

ventricular.combined.sct <- RunPCA(ventricular.combined.sct, verbose = FALSE)
ventricular.combined.sct <- RunUMAP(ventricular.combined.sct, reduction = "pca", dims = 1:40)

ventricular.combined.sct <- FindNeighbors(ventricular.combined.sct, reduction = "pca", dims = 1:40)
ventricular.combined.sct <- FindClusters(ventricular.combined.sct, resolution = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

Idents(object = ventricular.combined.sct) <- "integrated_snn_res.0.6"
DimPlot(ventricular.combined.sct,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/UMAP.res.0.6.D34.cells.5000.features.pdf", width = 10, height = 8)

#Segregation of clusters by sample
# UMAP of cells in each cluster by sample
DimPlot(ventricular.combined.sct, 
        label = TRUE, 
        split.by = "orig.ident")  + NoLegend()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/SplitSamples.UMAP.res.0.6.D34.cells.top.5000.features.pdf", width = 15, height = 5)

DimPlot(ventricular.combined.sct, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        #cols = DiscretePalette(12, palette = "stepped")[c(1,5,9,3,7,11,4,8,12)],
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/MergedSamples.UMAP.res.0.6.D34.cells.top.5000.features.pdf", width = 10, height = 8)

DimPlot(ventricular.combined.sct, 
        label = TRUE, 
        #cols = DiscretePalette(11, palette = "stepped")[c(1,5,9,2,6,10,3,7,11)],
        cols = custom_colors$discrete,
        repel = TRUE,
        group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/MergedSamples.UMAP.res.0.6.diff.colors.D34.cells.top.5000.features.pdf", width = 10, height = 8)
```



```{r}
# Count the number of clusters at different resolutions.
ventricular.combined.sct@meta.data <- select(ventricular.combined.sct@meta.data, -(11:33))
Ncluster <- c()
for(res in c(11:33)){
  Ncluster <- c(Ncluster, length(unique(as.vector(ventricular.combined.sct@meta.data[,res]))))
#unique(ventricular.cells@meta.data$SCT_snn_res.2)
}
res_ncluster <- data.frame(Res = c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0), Ncluster = Ncluster)
write.table(res_ncluster,"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/ResolutionVsNclusters.txt", quote = F, sep = "\t", row.names = F)
```


Segregation of clusters by cell cycle phase
```{r}
# Explore whether clusters segregate by cell cycle phase
DimPlot(ventricular.combined.sct,
        group.by = "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/Merged.CellCycle.UMAP.res.0.6.pdf", width = 10, height = 8)
DimPlot(ventricular.combined.sct,
        group.by = "Phase",
        split.by = "Phase")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/Split.CellCycle.UMAP.res.0.6.pdf", width = 25, height = 10)
```


Segregation of clusters by various sources of uninteresting variation
```{r}
# Determine metrics to plot present in seurat_integrated@meta.data
metrics <-  c("nCount_RNA", "nFeature_RNA", "S.Score", "G2M.Score", "percent.mt")

FeaturePlot(ventricular.combined.sct, 
            reduction = "umap", 
            features = metrics,
            pt.size = 0.4, 
            order = TRUE,
            #min.cutoff = 'q10',
            label = TRUE)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/Metrics.UMAP.res.0.6.pdf", width = 10, height = 10)
```


```{r}
# Save the object at this point so that it can easily be loaded back in without having to rerun the computationally intensive steps performed above, or easily shared with collaborators.
saveRDS(ventricular.combined.sct, file = "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/Clustered.ventricular.combined.sct.rds")

#ventricular.combined.sct <- readRDS("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/Clustered.ventricular.combined.sct.rds")
```

Exploring known cell type markers
```{r}
# Pacemaker cell genes
FeaturePlot(ventricular.combined.sct, 
            reduction = "umap", 
            features = c("TBX18", "SHOX2", "ISL1", "TBX3"), 
            order = FALSE,
            #min.cutoff = 'q10', 
            label = TRUE)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/PacemakerCell_Genes.UMAP.res.0.6.pdf", width = 12, height = 10)

# Violin plot
VlnPlot(ventricular.combined.sct, c("TBX18", "SHOX2", "ISL1", "TBX3"), group.by = "orig.ident", pt.size = 0, cols = custom_colors$discrete, ncol = 4)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/PacemakerCell_Genes.Violin.UMAP.res.0.6.pdf", width = 16, height = 5)
```

```{r}
DotPlot(ventricular.combined.sct, features = rev(c("NKX2-5", "TNNT2", "MYH6", "MYH7","TBX18", "SHOX2", "ISL1", "TBX3", "HCN4", "KCNJ3", "GJD3", "MYL2", "HEY2", "IRX4", "MYL7", "NPPA", "NR2F2", "MSX2", "TBX2")), cols= "Spectral", group.by = "orig.ident") + RotatedAxis() + coord_flip()
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/James_Markers_DotPlot.UMAP.res.0.6.pdf", width = 8, height = 6)
```

```{r}
# Single cell heatmap of feature expression
DoHeatmap(ventricular.combined.sct, features = c("NKX2-5", "TNNT2", "MYH6", "MYH7","TBX18", "SHOX2", "ISL1", "TBX3", "HCN4", "KCNJ3", "GJD3", "MYL2", "HEY2", "IRX4", "MYL7", "NPPA", "NR2F2", "MSX2", "TBX2"), size = 3, group.by = "orig.ident")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/James_Markers_Heatmap.UMAP.res.0.6.pdf")
```

Composition of samples and clusters
```{r}
table_samples_by_clusters <- ventricular.combined.sct@meta.data %>%
  group_by(orig.ident, integrated_snn_res.0.6) %>%
  summarize(count = n()) %>%
  spread(integrated_snn_res.0.6, count, fill = 0) %>%
  ungroup() %>%
  mutate(total_cell_count = rowSums(.[c(2:ncol(.))])) %>%
  dplyr::select(c('orig.ident', 'total_cell_count', everything())) %>%
  arrange(factor(orig.ident, levels = levels(ventricular.combined.sct@meta.data$orig.ident)))
write.table(as.data.frame(table_samples_by_clusters), "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_samples_by_clusters.txt", quote = F, sep = "\t")

table_clusters_by_samples <- ventricular.combined.sct@meta.data %>%
  dplyr::rename('cluster' = 'integrated_snn_res.0.6') %>%
  group_by(cluster, orig.ident) %>%
  summarize(count = n()) %>%
  spread(orig.ident, count, fill = 0) %>%
  ungroup() %>%
  mutate(total_cell_count = rowSums(.[c(2:ncol(.))])) %>%
  select(c('cluster', 'total_cell_count', everything())) %>%
  arrange(factor(cluster, levels = levels(ventricular.combined.sct@meta.data$integrated_snn_res.0.6)))
write.table(as.data.frame(table_clusters_by_samples), "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_clusters_by_samples.txt", quote = F, sep = "\t")

# Plot number of cells.
temp_labels <- ventricular.combined.sct@meta.data %>%
  group_by(orig.ident) %>%
  tally()

p1 <- table_samples_by_clusters %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'orig.ident') %>%
  mutate(sample = factor(orig.ident, levels = unique(ventricular.combined.sct@meta.data$orig.ident))) %>%
  ggplot(aes(sample, value)) +
  geom_bar(aes(fill = variable), position = 'stack', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = orig.ident, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 3, angle = 0
  ) +
  scale_fill_manual(name = 'Cluster', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Number of cells', labels = scales::comma, expand = c(0.01, 0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'left',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1, color = "black"),
    axis.text.y = element_text(color = "black"),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 0, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_samples_by_clusters_cell_number.pdf", p1, width = 8, height = 5)

temp_labels <- ventricular.combined.sct@meta.data %>%
  group_by(integrated_snn_res.0.6) %>%
  tally() %>%
  dplyr::rename('cluster' = integrated_snn_res.0.6)

p2 <- table_clusters_by_samples %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'cluster') %>%
  mutate(cluster = factor(cluster, levels = sort(unique(ventricular.combined.sct@meta.data$integrated_snn_res.0.6)))) %>%
  ggplot(aes(cluster, value)) +
  geom_bar(aes(fill = variable), position = 'stack', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = cluster, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Sample', values = custom_colors$discrete) +
  scale_y_continuous(labels = scales::comma, expand = c(0.01, 0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'right',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 10, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_clusters_by_samples_cell_number.pdf", p2, width = 10, height = 5)

# Plot percent of cells.
temp_labels <- ventricular.combined.sct@meta.data %>%
  group_by(orig.ident) %>%
  tally()

p1 <- table_samples_by_clusters %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'orig.ident') %>%
  mutate(sample = factor(orig.ident, levels = unique(ventricular.combined.sct@meta.data$orig.ident))) %>%
  ggplot(aes(sample, value)) +
  geom_bar(aes(fill = variable), position = 'fill', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = orig.ident, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Cluster', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Percentage [%]', labels = scales::percent_format(), expand = c(0.01,0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'left',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 0, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_samples_by_clusters_cell_percent.pdf", p1, width = 8, height = 5)

temp_labels <- ventricular.combined.sct@meta.data %>%
  group_by(integrated_snn_res.0.6) %>%
  tally() %>%
  dplyr::rename('cluster' = integrated_snn_res.0.6)

p2 <- table_clusters_by_samples %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'cluster') %>%
  mutate(cluster = factor(cluster, levels = sort(unique(ventricular.combined.sct@meta.data$integrated_snn_res.0.6)))) %>%
  ggplot(aes(cluster, value)) +
  geom_bar(aes(fill = variable), position = 'fill', stat = 'identity') +
  geom_text(
    data = temp_labels, aes(x = cluster, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Sample', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Percentage [%]', labels = scales::percent_format(), expand = c(0.01,0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'right',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 10, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_clusters_by_samples_cell_percent.pdf", p2, width = 10, height = 5)
```


```{r}
require(rBCS)
#ventricular.combined.sct@meta.data <- select(ventricular.combined.sct@meta.data, -seurat_clusters)
ventricular.combined.sct@meta.data <- mutate(ventricular.combined.sct@meta.data, seurat_clusters = ventricular.combined.sct@meta.data$integrated_snn_res.0.6)
ventricular.combined.sct@meta.data <- select(ventricular.combined.sct@meta.data, Sample = orig.ident, seurat_clusters, everything())
ExportSeurat(ventricular.combined.sct, "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/Clustered.ventricular.combined.sct.bcs")
```


Silhouette plot
```{r}
distance_matrix <- dist(Embeddings(ventricular.combined.sct[['pca']])[, 1:40])

sil_score_matrix <- matrix(NA, nrow=dim(ventricular.combined.sct@meta.data)[1], ncol=23)
for(i in 11:33){
  clusters <- ventricular.combined.sct@meta.data[,i]
  silhouette <- silhouette(as.numeric(clusters), dist = distance_matrix)
  sil_score_matrix[,i-10] <- silhouette[,3]
}

colnames(sil_score_matrix) = paste0("silhouette_score_res.", c(seq(0.01,0.09,0.01), 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0, 3.0))

ventricular.combined.sct@meta.data <- cbind(ventricular.combined.sct@meta.data, sil_score_matrix)

#ventricular.combined.sct@meta.data <- select(ventricular.combined.sct@meta.data, -35)

for(i in 34:56){
#for(i in 35){
  mean_silhouette_score <- mean(ventricular.combined.sct@meta.data[,i])
  tmp <- select(ventricular.combined.sct@meta.data, c(i-23, i))
  colnames(tmp) <- c("Cluster", "Sil_score")
  tmp <- mutate(tmp, barcode =  rownames(tmp))
  tmp <- arrange(tmp, Cluster, desc(Sil_score))
  tmp$barcode = factor(tmp$barcode, levels = tmp$barcode)
  p <- 
    ggplot(tmp) +
    #geom_col(aes(barcode, ventricular.combined.sct@meta.data[,i], fill = ventricular.combined.sct@meta.data)[,i-24], show.legend = TRUE) +
    geom_col(aes(barcode, Sil_score, fill = Cluster), show.legend = TRUE) +
    geom_hline(yintercept = mean_silhouette_score, color = 'red', linetype = 'dashed') +
    #scale_x_discrete(name = 'Cells') +
    xlab("Clusters") +
    ylab("Silhouette Score") +
    #scale_y_continuous(name = 'Silhouette score') +
    scale_fill_manual(values = custom_colors$discrete, name = 'Clusters') +
    theme_bw() +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
    ggsave(paste0("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/PCs/silhouette_plot_", names(ventricular.combined.sct@meta.data)[i], ".pdf"), p, height = 5, width = 8)
}

tmp_score <- select(ventricular.combined.sct@meta.data, 34:56)
library(reshape2)
long_tmp_score <- melt(tmp_score)

long_tmp_score$variable = str_replace(long_tmp_score$variable, "silhouette_score_res.", "")

library(ggpubr)
ggviolin(long_tmp_score, "variable", "value", fill = "variable",
   palette = custom_colors$discrete,
   add = "boxplot", add.params = list(fill = "white"), xlab = "Resolution", ylab = "Silhouette Score")
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/PCs/silhouette_violin_plot.pdf", width = 10, height = 5)


write.table(as.data.frame(colMeans(tmp_score)),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/PCs/ResolutionVsMean_Sil_Score.txt", quote = F, sep = "\t", row.names = T)
colMeans(tmp_score)
write.table(as.data.frame(apply(tmp_score,2,median)),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/PCs/ResolutionVsMedian_Sil_Score.txt", quote = F, sep = "\t", row.names = T)

```

Find markers at Res 0.6
```{r}
# find markers for every cluster compared to all remaining cells, report only the positive
# ones
ventricular.combined.sct.markers <- FindAllMarkers(ventricular.combined.sct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
save(ventricular.combined.sct.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.6.RData")

write.table(as.data.frame(ventricular.combined.sct.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.6.txt", quote = F, sep = "\t")

### This version of results were later move to "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/". The original folder "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/" was used for lenient criteria results. Jun 13 2022.
load("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.6.RData")

ls()
#[1] "filtered.samples.markers"

filtered.ventricular.samples.markers.fdr <- filter(ventricular.combined.sct.markers, p_val_adj < 0.01)

dim(filtered.ventricular.samples.markers.fdr)
#[1] 7587    7
dim(ventricular.combined.sct.markers)
#[1] 8811    7

filtered.ventricular.samples.markers.fdr %>%
    group_by(cluster) %>%
    top_n(n = 100, wt = avg_log2FC) -> top100

write.table(as.data.frame(top100),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.6_top100.txt", quote = F, sep = "\t")

####
#filtered.ventricular.samples.markers.fdr$cluster = as.vector(as.numeric(filtered.ventricular.samples.markers.fdr$cluster))
top100list <- matrix(0, nrow = 100, ncol = 20)
top100_df <- as.data.frame(top100)
for(i in seq(0,19,1)){
  j <- i+1
  top100list[1:length(top100_df[top100_df$cluster == i,7]),j] <- top100_df[top100_df$cluster == i,7]
}
colnames(top100list) <- paste0("C_", seq(0,19,1))

#top100list <- select(as.data.frame(top100list), c("C_18","C_6","C_8","C_3","C_15","C_13","C_21","C_12","C_5","C_2","C_10","C_9","C_16","C_26","C_27","C_19","C_20","C_1","C_24","C_23","C_25","C_17","C_22","C_14","C_4","C_7","C_11","C_0"))

write.table(as.data.frame(top100list),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.6_top100list.txt", quote = F, sep = "\t", row.names = F)
```


Find markers at Res 0.2
```{r}
# find markers for every cluster compared to all remaining cells, report only the positive
# ones
Idents(object = ventricular.combined.sct) <- "integrated_snn_res.0.2"
ventricular.combined.sct.markers <- FindAllMarkers(ventricular.combined.sct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
save(ventricular.combined.sct.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.2.RData")

write.table(as.data.frame(ventricular.combined.sct.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.2.txt", quote = F, sep = "\t")

filtered.ventricular.samples.markers.fdr <- filter(filtered.samples.markers, p_val_adj < 0.01)

dim(filtered.ventricular.samples.markers.fdr)
#[1] 25568     7
dim(filtered.samples.markers)
#[1] 26027     7

filtered.ventricular.samples.markers.fdr %>%
    group_by(cluster) %>%
    top_n(n = 100, wt = avg_log2FC) -> top100

write.table(as.data.frame(top100),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.2_top100.txt", quote = F, sep = "\t")

####
#filtered.ventricular.samples.markers.fdr$cluster = as.vector(as.numeric(filtered.ventricular.samples.markers.fdr$cluster))
top100list <- matrix(0, nrow = 100, ncol = 10)
top100_df <- as.data.frame(top100)
for(i in seq(0,9,1)){
  j <- i+1
  top100list[1:length(top100_df[top100_df$cluster == i,7]),j] <- top100_df[top100_df$cluster == i,7]
}
colnames(top100list) <- paste0("C_", seq(0,9,1))

#top100list <- select(as.data.frame(top100list), c("C_18","C_6","C_8","C_3","C_15","C_13","C_21","C_12","C_5","C_2","C_10","C_9","C_16","C_26","C_27","C_19","C_20","C_1","C_24","C_23","C_25","C_17","C_22","C_14","C_4","C_7","C_11","C_0"))

write.table(as.data.frame(top100list),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.2_top100list.txt", quote = F, sep = "\t", row.names = F)
```

```{r}
Idents(object = ventricular.combined.sct) <- "integrated_snn_res.0.4"
DimPlot(ventricular.combined.sct,
        reduction = "umap",
        label = TRUE,
        repel = TRUE,
        label.size = 6,
        cols = custom_colors$discrete)
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/UMAP.res.0.4.D34.cells.5000.features.pdf", width = 10, height = 8)

```

Composition of samples and clusters
```{r}
table_samples_by_clusters <- ventricular.combined.sct@meta.data %>%
  group_by(orig.ident, integrated_snn_res.0.4) %>%
  summarize(count = n()) %>%
  spread(integrated_snn_res.0.4, count, fill = 0) %>%
  ungroup() %>%
  mutate(total_cell_count = rowSums(.[c(2:ncol(.))])) %>%
  dplyr::select(c('orig.ident', 'total_cell_count', everything())) %>%
  arrange(factor(orig.ident, levels = levels(ventricular.combined.sct@meta.data$orig.ident)))
write.table(as.data.frame(table_samples_by_clusters), "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_samples_by_clusters_Res0.4.txt", quote = F, sep = "\t")

table_clusters_by_samples <- ventricular.combined.sct@meta.data %>%
  dplyr::rename('cluster' = 'integrated_snn_res.0.4') %>%
  group_by(cluster, orig.ident) %>%
  summarize(count = n()) %>%
  spread(orig.ident, count, fill = 0) %>%
  ungroup() %>%
  mutate(total_cell_count = rowSums(.[c(2:ncol(.))])) %>%
  select(c('cluster', 'total_cell_count', everything())) %>%
  arrange(factor(cluster, levels = levels(ventricular.combined.sct@meta.data$integrated_snn_res.0.4)))
write.table(as.data.frame(table_clusters_by_samples), "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_clusters_by_samples_Res0.4.txt", quote = F, sep = "\t")

# Plot number of cells.
temp_labels <- ventricular.combined.sct@meta.data %>%
  group_by(orig.ident) %>%
  tally()

p1 <- table_samples_by_clusters %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'orig.ident') %>%
  mutate(sample = factor(orig.ident, levels = unique(ventricular.combined.sct@meta.data$orig.ident))) %>%
  ggplot(aes(sample, value)) +
  geom_bar(aes(fill = variable), position = 'stack', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = orig.ident, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 3, angle = 0
  ) +
  scale_fill_manual(name = 'Cluster', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Number of cells', labels = scales::comma, expand = c(0.01, 0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'left',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1, color = "black"),
    axis.text.y = element_text(color = "black"),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 0, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_samples_by_clusters_cell_number_Res0.4.pdf", p1, width = 8, height = 5)

temp_labels <- ventricular.combined.sct@meta.data %>%
  group_by(integrated_snn_res.0.4) %>%
  tally() %>%
  dplyr::rename('cluster' = integrated_snn_res.0.4)

p2 <- table_clusters_by_samples %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'cluster') %>%
  mutate(cluster = factor(cluster, levels = sort(unique(ventricular.combined.sct@meta.data$integrated_snn_res.0.4)))) %>%
  ggplot(aes(cluster, value)) +
  geom_bar(aes(fill = variable), position = 'stack', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = cluster, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Sample', values = custom_colors$discrete) +
  scale_y_continuous(labels = scales::comma, expand = c(0.01, 0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'right',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 10, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_clusters_by_samples_cell_number_Res0.4.pdf", p2, width = 10, height = 5)

# Plot percent of cells.
temp_labels <- ventricular.combined.sct@meta.data %>%
  group_by(orig.ident) %>%
  tally()

p1 <- table_samples_by_clusters %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'orig.ident') %>%
  mutate(sample = factor(orig.ident, levels = unique(ventricular.combined.sct@meta.data$orig.ident))) %>%
  ggplot(aes(sample, value)) +
  geom_bar(aes(fill = variable), position = 'fill', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = orig.ident, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Cluster', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Percentage [%]', labels = scales::percent_format(), expand = c(0.01,0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'left',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 0, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_samples_by_clusters_cell_percent_Res0.4.pdf", p1, width = 8, height = 5)

temp_labels <- ventricular.combined.sct@meta.data %>%
  group_by(integrated_snn_res.0.4) %>%
  tally() %>%
  dplyr::rename('cluster' = integrated_snn_res.0.4)

p2 <- table_clusters_by_samples %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'cluster') %>%
  mutate(cluster = factor(cluster, levels = sort(unique(ventricular.combined.sct@meta.data$integrated_snn_res.0.4)))) %>%
  ggplot(aes(cluster, value)) +
  geom_bar(aes(fill = variable), position = 'fill', stat = 'identity') +
  geom_text(
    data = temp_labels, aes(x = cluster, y = Inf, label = paste0('', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Sample', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Percentage [%]', labels = scales::percent_format(), expand = c(0.01,0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'right',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 10, unit = 'pt')
  )
ggsave("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/table_clusters_by_samples_cell_percent_Res0.4.pdf", p2, width = 10, height = 5)
```

Find markers at Res 0.4
```{r}
# find markers for every cluster compared to all remaining cells, report only the positive
# ones
ventricular.combined.sct.markers <- FindAllMarkers(ventricular.combined.sct, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
save(ventricular.combined.sct.markers, file="/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.4.RData")

write.table(as.data.frame(ventricular.combined.sct.markers),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.4.txt", quote = F, sep = "\t")

### This version of results were later move to "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/". The original folder "/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/" was used for lenient criteria results. Jun 13 2022.
#load("/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.4.RData")

ls()
#[1] "filtered.samples.markers"

filtered.ventricular.samples.markers.fdr <- filter(ventricular.combined.sct.markers, p_val_adj < 0.01)

table(filtered.ventricular.samples.markers.fdr$cluster)
#   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14 
# 162  159  366  681   86  200  698  198  335 1320  183  818  692  130  385

filtered.ventricular.samples.markers.fdr.log2fc <- filter(ventricular.combined.sct.markers, p_val_adj < 0.01 & avg_log2FC > log2(1.5))

table(filtered.ventricular.samples.markers.fdr.log2fc$cluster)
#   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14 
# 133  130  317  643   59  180  649  160  300 1265  174  792  678  121  382 

dim(filtered.ventricular.samples.markers.fdr.log2fc)
#[1] 5983    7
dim(filtered.ventricular.samples.markers.fdr)
#[1] 6413    7
dim(ventricular.combined.sct.markers)
#[1] 7257    7

filtered.ventricular.samples.markers.fdr.log2fc %>%
    group_by(cluster) %>%
    slice_max(n = 100, order_by = -p_val_adj) -> top100

filtered.ventricular.samples.markers.fdr.log2fc %>%
    group_by(cluster) %>%
    slice_head(n = 100) -> top100

write.table(as.data.frame(top100),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.4_top100_fdr.txt", quote = F, sep = "\t")
####
#filtered.pacemaker.samples.markers.fdr$cluster = as.vector(as.numeric(filtered.pacemaker.samples.markers.fdr$cluster))
top100list <- matrix(0, nrow = 100, ncol = 15)
top100_df <- as.data.frame(top100)
for(i in seq(0,14,1)){
  j <- i+1
  top100list[1:length(top100_df[top100_df$cluster == i,7]),j] <- top100_df[top100_df$cluster == i,7]
}
colnames(top100list) <- paste0("C_", seq(0,14,1))

#top100list <- select(as.data.frame(top100list), c("C_18","C_6","C_8","C_3","C_15","C_13","C_21","C_12","C_5","C_2","C_10","C_9","C_16","C_26","C_27","C_19","C_20","C_1","C_24","C_23","C_25","C_17","C_22","C_14","C_4","C_7","C_11","C_0"))

write.table(as.data.frame(top100list),"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/FindAllMarkers_Res0.4_top100list.txt", quote = F, sep = "\t", row.names = F)
```

```{r}
cellnumber <-  as.data.frame(table(ventricular.combined.sct@meta.data$integrated_snn_res.0.4))
names(cellnumber) <- c("Cluster","CellNumber")
write.table(cellnumber,"/Data/iPSC_pacemaker/Seurat/D34_First/VentricularCells/Integration/UMAP/Res0.4_CellNumber.txt", quote = F, sep = "\t", row.names = F)
```








